### 2020-08-21 sections/expertise.tex
Create expertise.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Expertise}

\begin{cvskills}

\cvskill
{Interests}
{Cabinets composition and allocation, ministerial turnover, political elites socialisation, civil service}

\cvskill
{}
{systems, political participation, digital media, electronic government}\\

\cvskill
{Techniques} 
{Survival analysis, regression modelling, multilevel models, factor analysis, spatial analysis, social}

\cvskill
{} 
{network analysis, qualitative comparative analysis, topic modelling} \\

\end{cvskills}