### 2020-08-21 sections/papers-isi.tex
Create papers-isi.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Peer-Reviewed Articles}

\cvsubsection{Social Sciences Citation Index}

\begin{cventries}

\begin{benumerate}{5}
\item{\small Gonz\'alez-Bustamante, B. ({\itshape forthcoming}). Evolution and Early Government Responses to COVID-19 in South America. {\itshape World Development}. {\footnotesize \scshape [{\bfseries 3.869} 2019-JIF, {\bfseries Q1} Economics]}.}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., Carvajal, A., \& Gonz\'alez, A. (2020). Determinants of E-Government in the Municipalities: Evidence from the Chilean Case. {\itshape Gesti\'on y Pol\'itica P\'ublica, XXIX}(1), 97--129. {\footnotesize \scshape [in Spanish, {\bfseries 0.222} 2019-JIF, {\bfseries Q4} Public Administration]}. {\scshape doi:} \href{http://dx.doi.org/10.29265/gypp.v29i1.658}{\textcolor{blue}{10.29265/gypp.v29i1.658}}. {\scshape {\footnotesize SocArXiv} doi}: \href{https://doi.org/10.31235/osf.io/fze3x}{\textcolor{blue}{10.31235/osf.io/fze3x}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2019). The Politics-Administration Dichotomy: A Case Study of the Chilean Executive During the Democratic Post-Transition. {\itshape Bulletin of Latin American Research}. Early View. {\footnotesize \scshape [{\bfseries 0.586} 2019-JIF, {\bfseries Q3} Area Studies]}. {\scshape doi}: \href{https://doi.org/10.1111/blar.13044}{\textcolor{blue}{10.1111/blar.13044}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2019). Gaps, Representation, and Elite-Mass Congruence in Chile and Uruguay. {\itshape Convergencia. Revista de Ciencias Sociales}, (80), 1--27. {\footnotesize \scshape [in Spanish, {\bfseries 0.704} 2019-JIF, {\bfseries Q4} Sociology]}. \\ {\scshape doi}: \href{https://doi.org/10.29101/crcs.v26i80.11097}{\textcolor{blue}{10.29101/crcs.v26i80.11097}}. {\scshape {\footnotesize SocArXiv} doi}: \href{https://doi.org/10.31235/osf.io/cqym8}{\textcolor{blue}{10.31235/osf.io/cqym8}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Garrido-Vergara, L. (2018). Socialization, Trajectories and Post-Career of Ministers in Chile, 1990-2010. {\itshape Pol\'itica y Gobierno, XXV}(1), 31--64. {\footnotesize \scshape [in Spanish, {\bfseries 0.714} 2019-JIF, {\bfseries Q4} Political Science]}. {\scshape url}: \href{http://www.politicaygobierno.cide.edu/index.php/pyg/article/view/1080}{\textcolor{blue}{www.politicaygobierno.cide.edu}}}\vspace{1mm}
\end{benumerate}

\end{cventries}