### 2020-08-21 sections/papers-scopus.tex
Create papers-scopus.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Peer-Reviewed Articles}

\cvsubsection{Scopus \& Emerging Sources Citation Index}

\begin{cventries}

\begin{benumerate}{7}
\item{\small Olivares, A. {\itshape et al.} (2020). New Challenges, Approaches, and Perspectives in the Study of Political Elites. {\itshape Iberoamericana, XX}(74), 229--259. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{https://journals.iai.spk-berlin.de/index.php/iberoamericana/article/view/2736}{\textcolor{blue}{https://journals.iai.spk-berlin.de}}. \\ {\scshape {\footnotesize SocArXiv} doi}: \href{https://doi.org/10.31235/osf.io/syqu4}{\textcolor{blue}{10.31235/osf.io/syqu4}}}\vspace{1mm}

\item{\small Maillet, A., Gonz\'alez-Bustamante, B., \& Olivares, A. (2019). Public-Private Circulation and the Revolving Door in the Chilean Executive Branch (2000-2014). {\itshape Latin American Business Review, 20}(4), 367--387. {\scshape doi}: \href{https://doi.org/10.1080/10978526.2019.1652099}{\textcolor{blue}{10.1080/10978526.2019.1652099}}}\vspace{1mm}

\item{\small Barr\'ia, D., Gonz\'alez-Bustamante, B., \& Cisternas, C. (2019). The Literature on Open Government in Spanish. Analysis of the Dynamics of Production and Citation. {\itshape N\'oesis. Revista de Ciencias Sociales y Humanidades, 28}(56), 22--42. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{http://dx.doi.org/10.20983/noesis.2019.2.3}{\textcolor{blue}{10.20983/noesis.2019.2.3}}}\vspace{1mm}

\item{\small Del Valle, N., \& Gonz\'alez-Bustamante, B. (2018). Political Agenda, Journalism and Digital Media in Chile: Research Notes on Media Pluralism. {\itshape Perspectivas de la Comunicaci\'on, 11}(1), 291--326. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{http://revistas.ufro.cl/ojs/index.php/perspectivas/article/view/1146}{\textcolor{blue}{http://revistas.ufro.cl}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Olivares, A. (2016). Cabinet Turnover and Survival of Ministers in Chile During the Concertacion Governments (1990-2010). {\itshape Colombia Internacional}, (87), 81--108. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{https://doi.org/10.7440/colombiaint87.2016.04}{\textcolor{blue}{10.7440/colombiaint87.2016.04}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. {\itshape et al.} (2016). Civil Service in Chile, Analysis of Senior Civil Servants on the Top Rung of the Hierarchy (2003-13). {\itshape Revista de Administra\c{c}\~ao P\'ublica, 50}(1), 59--79. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{http://dx.doi.org/10.1590/0034-7612145767}{\textcolor{blue}{10.1590/0034-7612145767}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Olivares, A. (2015). Undersecretaries Rotation in Chile: An Exploration of the Second Government Line, 1990-2014. {\itshape Revista de Gesti\'on P\'ublica, IV}(2), 151--190. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{https://www.researchgate.net/publication/321977889_Rotacion_de_subsecretarios_en_Chile_una_exploracion_de_la_segunda_linea_gubernamental_1990-2014}{\textcolor{blue}{www.researchgate.net}}} \vspace{1mm}
\end{benumerate}

\end{cventries}