### 2020-08-21 sections/papers-scielo.tex
Create papers-scielo.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Peer-Reviewed Articles}

\cvsubsection{ SciELO \& Latindex}

\begin{cventries}

\begin{benumerate}{9}
\item{\small Gonz\'alez-Bustamante, B. (2016). Political, Economic, and Intellectual Elites: A Growing Research Agenda for Political Science. {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 54}(1), 7--17. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/article/view/42690}{\textcolor{blue}{https://revistapolitica.uchile.cl}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Cisternas, C. (2016). Political Elites in the Chilean Legislative Branch: The Chamber of Deputies (1990-2014). {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 54}(1), 19--52. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/article/view/42691}{\textcolor{blue}{https://revistapolitica.uchile.cl}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2015). Assessing Twitter as a Public Opinion Measure: An Exploration of the Bachelet's Arrival to the Chilean Presidential Campaign 2013. {\itshape Revista SAAP, 9}(1), 119--141. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{http://ref.scielo.org/dwzhns}{\textcolor{blue}{http://ref.scielo.org}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2015). Electoral Success and Campaign Spending in Elections of Senators and Deputies in Chile 2013. {\itshape Pol\'iticas P\'ublicas, 8}(1), 21--35. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{http://www.revistas.usach.cl/ojs/index.php/politicas/article/view/2182}{\textcolor{blue}{www.revistas.usach.cl}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2014). Popular Election of Regional Councillors 2013. Performance of Political, Family, and Economic Capital in a New Electoral Arena in Chile. {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 52}(2), 49--91. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/article/view/36137}{\textcolor{blue}{https://revistapolitica.uchile.cl}}} \vspace{1mm}

\item{\small Olivares, A. {\itshape et al.} (2014). Think Tanks in the Cabinet: An Exploration of the Chilean Case (2006-2014). {\itshape Revista de Sociolog\'ia,} (29), 37--54. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{https://revistadesociologia.uchile.cl/index.php/RDS/article/view/36177}{\textcolor{blue}{https://revistadesociologia.uchile.cl}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2013). Factors Affecting Access and Permanence of the Governmental Political Elite in Chile (1990-2010). {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 51}(1), 119--153. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/article/view/27436}{\textcolor{blue}{https://revistapolitica.uchile.cl}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2013). The Study of the Elites in Chile: Theoretical and Methodological Approaches. {\itshape Intersticios Sociales,} (6), 1--20. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{http://ref.scielo.org/zrnp2k}{\textcolor{blue}{http://ref.scielo.org}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Henr\'iquez, G. (2012). Digital Campaigns: Branding or Political Participation? {\itshape M\'as Poder Local,} (12), 32--39. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{https://www.researchgate.net/publication/260517478_Campanas_digitales_Branding_o_participacion_politica_El_rol_de_las_redes_sociales_en_la_ultima_campana_presidencial_chilena}{\textcolor{blue}{www.researchgate.net}}} \vspace{1mm}
\end{benumerate}

\end{cventries}