### 2020-08-21 sections/projects.tex
Create projects.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% DEPRECATED

\cvsection{Funding \& Research Projects}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Chilean Political Science Impact Ranking} 
{This research have not received any specific grant. {\scshape doi:} \href{http://doi.org/10.17605/OSF.IO/C8PRA}{\textcolor{blue}{10.17605/OSF.IO/C8PRA}}} 
{Coordinator}
{2020}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{COVID-19 Pandemic in South America}
{This research have not received any specific grant. {\scshape doi:} \href{http://doi.org/10.17605/OSF.IO/6FM7X}{\textcolor{blue}{10.17605/OSF.IO/6FM7X}}}
{Coordinator}
{2020}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Gender Turnout and Elected Representatives in Iceland Municipal Elections}
{Prof. Petra Schleiter, Department of Politics and International Relations and St Hilda's College, University of Oxford}
{Research Assistant}
{2020}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Survival of the Senior Civil Servants in the Chilean Executive Branch}
{Funded by the Universidad de Santiago de Chile (Academic Performance Agreement). {\scshape doi:} \href{http://doi.org/10.17605/OSF.IO/WBF6M}{10.17605/OSF.IO/WBF6M}}
{Researcher}
{2018}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Political System and Visibility of Actors in the Public Debate Articulated by the Elite Press, 1990-2018}
{Funded by the Universidad Alberto Hurtado (XVI Promotion of Research Fund, Ximena Orchard, Principal Investigator)}
{Researcher}
{2019}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Transparency and Publication Dynamics on the Latin American Public Administration}
{Funded by the Universidad de Santiago (USA1498.57, Diego Barría, Principal Investigator)}
{Researcher}
{2017}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Public Service Motivation, Open Government and Preferences on Public Services}
{Funded by the Universidad de Santiago (USA1498.37, Diego Barría, Principal Investigator)}
{Researcher}
{2016}
\end{cvhonors}

\vspace{1mm}

%% PENDING PROJECTS