### 2020-08-21 spanish/sections/references.tex
Create references.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in Spanish
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Referencias}

\begin{publications}

\begin{itemize}
\item[]{\small {\bfseries Petra Schleiter}, Profesora en Política Comparada, Universidad de Oxford, RU}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:petra.schleiter@politics.ox.ac.uk}{\textcolor{blue}{petra.schleiter@politics.ox.ac.uk}}}} \vspace{1.5mm}
\item[]{\small {\bfseries Gonzalo Delamaza}, Profesor Titular, Universidad de Los Lagos, Chile}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:gonzalo.delamaza@ulagos.cl}{\textcolor{blue}{gonzalo.delamaza@ulagos.cl}}}} \vspace{1.5mm}
\item[]{\small {\bfseries Miguel \'Angel L\'opez}, Profesor Asistente, Universidad de Chile, Chile}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:mlopez@uchile.cl}{\textcolor{blue}{mlopez@uchile.cl}}}} \pagebreak %% \vspace{1.5mm}
\item[]{\small {\bfseries Alfredo Joignant}, Profesor Titular, Universidad Diego Portales, Chile}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:alfredo.joignant@mail.udp.cl}{\textcolor{blue}{alfredo.joignant@mail.udp.cl}}}} \vspace{1.5mm}
\item[]{\small {\bfseries Jaime Baeza}, Profesor Asistente, Universidad de Chile, Chile}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:jbaezaf@u.uchile.cl}{\textcolor{blue}{jbaezaf@u.uchile.cl}}}} \vspace{1.5mm}
\item[]{\small {\bfseries Marcel Thez\'a}, Profesor Asociado, Universidad de Los Lagos, Chile}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:marcel.theza@ulagos.cl}{\textcolor{blue}{marcel.theza@ulagos.cl}}}} \vspace{1.5mm}
\end{itemize}

\end{publications}