### 2020-08-21 spanish/sections/expertise.tex
Update expertise.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in Spanish
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Especialidades}

\begin{cvskills}

\cvskill
{Interests}
{Composición y nombramiento de gabinetes, rotación ministerial, socialización de élites políticas,}

\cvskill
{}
{sistemas de servicio civil, participación política, medios digitales, gobierno electrónico}\\

\cvskill
{Techniques} 
{Survival analysis, regression modelling, multilevel models, factor analysis, spatial analysis, social}

\cvskill
{} 
{network analysis, qualitative comparative analysis, topic modelling} \\

\end{cvskills}