### 2020-08-21 spanish/sections/isi-papers.tex
Update isi-papers.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in Spanish
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Publicaciones revisadas por pares}

\cvsubsection{Social Sciences Citation Index (SCCI/WoS-ISI)}

\begin{publications}

\begin{benumerate}{5}
\item{\small Gonz\'alez-Bustamante, B. ({\itshape forthcoming}). Evolution and Early Government Responses to COVID-19 in South America. {\itshape World Development}.}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., Carvajal, A., \& Gonz\'alez, A. (2020). Determinantes del gobierno electrónico en las municipalidades: Evidencia del caso chileno. {\itshape Gesti\'on y Pol\'itica P\'ublica, XXIX}(1), 97--129. \\ {\scshape doi:} \href{http://dx.doi.org/10.29265/gypp.v29i1.658}{\textcolor{blue}{10.29265/gypp.v29i1.658}}. {\scshape {\footnotesize SocArXiv} doi}: \href{https://doi.org/10.31235/osf.io/fze3x}{\textcolor{blue}{10.31235/osf.io/fze3x}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2019). The Politics-Administration Dichotomy: A Case Study of the Chilean Executive During the Democratic Post-Transition. {\itshape Bulletin of Latin American Research}. Early View. \\ {\scshape doi}: \href{https://doi.org/10.1111/blar.13044}{\textcolor{blue}{10.1111/blar.13044}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2019). Brechas, representación y congruencia élite-ciudadanía en Chile y Uruguay. {\itshape Convergencia. Revista de Ciencias Sociales}, (80), 1--27. {\scshape doi}: \href{https://doi.org/10.29101/crcs.v26i80.11097}{\textcolor{blue}{10.29101/crcs.v26i80.11097}}. \\ {\scshape {\footnotesize SocArXiv} doi}: \href{https://doi.org/10.31235/osf.io/cqym8}{\textcolor{blue}{10.31235/osf.io/cqym8}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Garrido-Vergara, L. (2018). Socialización, trayectorias y poscarrera de ministros en Chile, 1990-2010. {\itshape Pol\'itica y Gobierno, XXV}(1), 31--64. {\scshape url}: \href{http://www.politicaygobierno.cide.edu/index.php/pyg/article/view/1080}{\textcolor{blue}{www.politicaygobierno.cide.edu}}}\vspace{1mm}
\end{benumerate}

\end{publications}