### 2020-08-23 spanish/sections/languages.tex
Update languages.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in Spanish
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Idiomas y manejo de softwares}

\begin{cvskills}

\cvskill
{Idiomas} 
{Español (hablante nativo), inglés (fluido), portugués (habilidades de lectura avanzadas),}

\cvskill
{} 
{francés (habilidades de lectura elementales)} \\

\cvskill
{Programming} 
{R, Python, Stata, {\scshape spss/pspp}, {\scshape ucinet}, NetDraw, RSiena, Gephi, Tosmana, Fuzzy-Sets, ArcGIS,}

\cvskill
{} 
{{\scshape atlas}.ti, EndNote, Zotero, {\scshape HPoP}, {\LaTeX}, Bib{\TeX}, JabRef, MS Office, WordPress, Moodle, {\scshape html},}

\cvskill
{} 
{{\scshape css}, {\scshape xhtml}, Notepad{\footnotesize ++}} \\

\end{cvskills}