### 2021-04-26 sections/languages.tex
Update languages.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Languages \& Software}

\begin{cvskills}

\cvskill
{Languages} 
{Spanish (native speaker), English (fluent), Portuguese (advanced reading skills), French}

\cvskill
{} 
{(elemental reading skills)} \\

\cvskill
{Programming \&} 
{R, Python, Stata, {\scshape spss/pspp}, {\scshape ucinet}, NetDraw, RSiena, Gephi, Tosmana, Fuzzy-Sets, ArcGIS,}

\cvskill
{Software} 
{{\scshape atlas}.ti, EndNote, Zotero, {\scshape HPoP}, {\LaTeX}, Bib{\TeX}, JabRef, MS Office, WordPress, Moodle, {\scshape html},}

\cvskill
{} 
{{\scshape css}, {\scshape xhtml}} \\

\end{cvskills}