### 2021-09-23 sections/selected-training.tex
Update selected-training.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.txt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Selected Certifications \& Training}

%% OX|BER
%% USACH-Nuffield
%% R Conference

\begin{cvhonors}
\cvconf
{Research Data Workshop IT Centre at the University of Oxford} 
{Oxford, UK}
{2019}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Longitudinal Networks with RSiena Workshop at the P Universidad Cat\'olica}
{Santiago, Chile}
{2018}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Universidad de Santiago Winter School for Academics}
{Santiago, Chile}
{2016}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Nuffield-USACH CEES Experimental Summer School}
{Santiago, Chile}
{2016}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Universidad de Santiago Winter School for Academics}
{Santiago, Chile}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Latin American Political Science Association Methodological School}
{Lima, Peru}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Universidad Cat\'olica de Uruguay Winter School in Methods and Data Analysis}
{MVD, Uruguay}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{P Universidad Ca\'olica Summer School in Mixed Methods}
{Santiago, Chile}
{2014}
\end{cvhonors}

\vspace{1mm}