### 2021-11-24 sections/awards.tex
Update awards.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Honours \& Awards}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{St Hilda's College Travel Grant} 
{Allowances for conference participation in the USA, {\pounds} 950} 
{Oxford, UK}
{2022}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{St Hilda's College Muriel Wise Fund and Book Grant} 
{Research grant for the training of machine learning algorithms, DPhil (PhD) project data storage, and allowances for books, {\pounds} 530} 
{Oxford, UK}
{2021}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{St Hilda's College Travel and Book Grants} 
{Allowances for conference participation in Portugal and books (travel grant was declined because of coronavirus pandemic), {\pounds} 745} 
{Oxford, UK}
{2020}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Becas Chile Doctoral Scholarship Abroad} 
{Tuition fees, transport and living expenses for pursuing a DPhil (PhD) at the University of Oxford (ANID/PFCHA/72200340), {\pounds} 160,000}
{Oxford, UK}
{2019}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{LASA Travel Grant} 
{Travel support for the XXXIV International Congress of the Latin American Studies Association (LASA/40071), {\pounds} 400}
{New York, USA}
{2016}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{CESS Summer School Scholarship}
{Financial support for the Experimental Summer School Methods at the Nuffield-USACH Centre for Experimental Social Sciences (CESS), {\pounds} 200}
{Santiago, Chile}
{2016}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{ALACIP Methodological School Prize}
{Best methodological work at the Latin American Political Association School (ALACIP)}
{Lima, Peru}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{ACCP Winter School Scholarship}
{Financial support for the Winter School in Methods and Data Analysis at the Universidad Cat\'olica de Uruguay, awarded by the Chilean Political Science Association (ACCP), {\pounds} 300}
{Montevideo, URY}
{ 2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{PUC Summer School Scholarship}
{Financial support for the Summer School in Mixed Methods at the Pontificia Universidad Cat\'olica de Chile (PUC), {\pounds} 170}
{Santiago, Chile}
{2014} 
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Comunicracia Prize}
{Best paper of the political communication area at the XI Argentinian Congress of Political Science (double-blind peer-review), {\pounds} 170}
{Parana, Argentina}
{2013}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Highest Rating MA Dissertation}
{The highest rating and the congratulations from the jury at the Universidad de Chile}
{Santiago, Chile}
{2013}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{FONDECYT Thesis Writing Grant}
{Financial support for MA field work (Prof. Gonzalo Delamaza, Supervisor, ANID/FONDECYT/1100877), {\pounds} 1,150}
{Santiago, Chile}
{2012}
\end{cvhonors}

\vspace{1mm}