### 2021-11-24 spanish/sections/experience.tex
Update experience.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in Spanish
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Experiencia profesional relevante}

\begin{cventries}
\cventry
{Profesor Instructor} 
{Universidad de Santiago de Chile} 
{Santiago, Chile} 
{2019 - en comisión}
{Departamento de Gestión y Políticas Públicas, Facultad de Administración y Economía}\vspace{1.5mm}

\cventry
{Profesor Adjunto} 
{Universidad de Santiago de Chile} 
{Santiago, Chile} 
{2014 - 2018}
{Departamento de Gestión y Políticas Públicas, Facultad de Administración y Economía}\vspace{1.5mm}

\cventry
{Consultor} 
{Programa de las Naciones Unidas para el Desarrollo (PNUD)} 
{Santiago, Chile} 
{2016 - 2017}
{Programa de Gobernabilidad Democrática} \vspace{1.5mm}

\cventry
{Docente} 
{Universidad de Chile} 
{Santiago, Chile} 
{2013 - 2016}
{Escuela de Gobierno y Gestión Pública, Instituto de Asuntos Públicos}\vspace{1.5mm}

\end{cventries}