### 2021-11-24 spanish/sections/editor.tex
Update editor.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in Spanish
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Números especiales editados}

\begin{publications}

\begin{benumerate}{2}
\item{\small Gonz\'alez-Bustamante, B., \& Luci, F. (2022, {\itshape próximamente}). Political Elites in Latin America: Socialisation, Trajectories, and Capitals. {\itshape Pléyade}.}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2016). Élites políticas, económicas e intelectuales. {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 54}(1). {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/issue/view/4365}{\textcolor{blue}{https://revistapolitica.uchile.cl}}} \vspace{1mm}
\end{benumerate}

\end{publications}