### 2021-11-24 sections/reviewer.tex
Update reviewer.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Peer Reviewer \& Expert}

\begin{publications}

\begin{itemize}
\item{\small Varieties of Democracy Project, University of Gothenburg, Sweden, 2014, 2018-2020}
\item{\small Global Academic Reputation Survey, Clarivate Analytics, USA, 2020}
\item{\small Sectoral Commission for Research, Universidad de la Rep\'ublica, Uruguay, 2016, 2018}
\item{\small Times Higher Education World University Rankings, {\scshape Elsevier}, 
Netherlands, 2018-2019}
\item{\small Reviewer in {\scshape WoS-ISI/SSCI/ESCI} \& {\scshape Scopus} journals, since 2013}
\item{\small Reviewer in {\scshape SciELO} \& {\scshape Latindex} journals, since 2013}
\end{itemize}

\vspace{1mm}
\end{publications}