### 2021-11-24 sections/experience.tex
Update experience.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Relevant Employment}

\begin{cventries}
\cventry
{Instructor Professor} 
{Universidad de Santiago de Chile} 
{Santiago, Chile} 
{2019 - on leave}
{Management and Public Policies Department, Faculty of Administration and Economics} \vspace{1.5mm}

\cventry
{Adjunct Professor} 
{Universidad de Santiago de Chile} 
{Santiago, Chile} 
{2014 - 2018}
{Management and Public Policies Department, Faculty of Administration and Economics} \vspace{1.5mm}

\cventry
{Consultant} 
{United Nations Development Programme (UNDP)} 
{Santiago, Chile} 
{2016 - 2017}
{Democratic Governance Programme} \vspace{1.5mm}

\cventry
{Lecturer} 
{Universidad de Chile} 
{Santiago, Chile} 
{2013 - 2016}
{School of Government and Public Management, Institute of Public Affairs}\vspace{1.5mm}

\end{cventries}