### 2021-11-25 sections/papers.tex
Update papers.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Peer-Reviewed Articles}

\cvsubsection{Social Sciences Citation Index (WoS-ISI/SCCI)}

\begin{publications}

\begin{benumerate}{5}

\item{\small Gonz\'alez-Bustamante, B. (2021). Evolution and Early Government Responses to COVID-19 in South America. {\itshape World Development, 137}, 105180. {\scshape doi:} \href{https://doi.org/10.1016/j.worlddev.2020.105180}{\textcolor{blue}{10.1016/j.worlddev.2020.105180}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2020). The Politics-Administration Dichotomy: A Case Study of the Chilean Executive During the Democratic Post-Transition. {\itshape Bulletin of Latin American Research, 39}(5), 582--597. {\scshape doi}: \href{https://doi.org/10.1111/blar.13044}{\textcolor{blue}{10.1111/blar.13044}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., Carvajal, A., \& Gonz\'alez, A. (2020). Determinants of E-Government in the Municipalities: Evidence from the Chilean Case. {\itshape Gesti\'on y Pol\'itica P\'ublica, XXIX}(1), 97--129. {\footnotesize \scshape [in Spanish]}. {\scshape doi:} \href{http://dx.doi.org/10.29265/gypp.v29i1.658}{\textcolor{blue}{10.29265/gypp.v29i1.658}}. {\scshape \footnotesize SocArXiv}: \href{https://doi.org/10.31235/osf.io/fze3x}{\textcolor{blue}{10.31235/osf.io/fze3x}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2019). Gaps, Representation, and Elite-Mass Congruence in Chile and Uruguay. {\itshape Convergencia. Revista de Ciencias Sociales}, 80, 1--27. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{https://doi.org/10.29101/crcs.v26i80.11097}{\textcolor{blue}{10.29101/crcs.v26i80.11097}}. {\scshape url}:  \href{https://www.redalyc.org/jatsRepo/105/10559568002/index.html}{\textcolor{blue}{www.redalyc.org}}. {\scshape \footnotesize SocArXiv}: \href{https://doi.org/10.31235/osf.io/cqym8}{\textcolor{blue}{10.31235/osf.io/cqym8}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Garrido-Vergara, L. (2018). Socialization, Trajectories and Post-Career of Ministers in Chile, 1990-2010. {\itshape Pol\'itica y Gobierno, XXV}(1), 31--64. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{http://www.politicaygobierno.cide.edu/index.php/pyg/article/view/1080}{\textcolor{blue}{www.politicaygobierno.cide.edu}}}\vspace{1mm}

\end{benumerate}

\end{publications}

\cvsubsection{Scopus \& Emerging Sources Citation Index (WoS-ISI/ESCI)}

\begin{publications}

\begin{benumerate}{8}

\item{\small González-Bustamante, B., Astete, M., \& Orvenes, B. (2020). Senior Public Managers: A Novel Dataset on Members of the Chilean Civil Service. {\itshape Revista de Gestión Pública, IX}(2), 151--169. {\footnotesize \scshape [in Spanish and English]}. {\scshape doi}: \href{https://doi.org/10.22370/rgp.2020.9.2.2920}{\textcolor{blue}{10.22370/rgp.2020.9.2.2920}}. {\scshape \footnotesize SocArXiv}: \href{https://doi.org/10.31235/osf.io/vshcz}{\textcolor{blue}{10.31235/osf.io/vshcz}}}\vspace{1mm}

\item{\small Olivares, A. {\itshape et al.} (2020). New Challenges, Approaches, and Perspectives in the Study of Political Elites. {\itshape Iberoamericana, XX}(74), 229--259. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{https://doi.org/10.18441/ibam.20.2020.74.229-259}{\textcolor{blue}{10.18441/ibam.20.2020.74.229-259}}. \\{\scshape \footnotesize SocArXiv}: \href{https://doi.org/10.31235/osf.io/syqu4}{\textcolor{blue}{10.31235/osf.io/syqu4}}}\vspace{1mm}

\item{\small Maillet, A., Gonz\'alez-Bustamante, B., \& Olivares, A. (2019). Public-Private Circulation and the Revolving Door in the Chilean Executive Branch (2000-2014). {\itshape Latin American Business Review, 20}(4), 367--387. {\scshape doi}: \href{https://doi.org/10.1080/10978526.2019.1652099}{\textcolor{blue}{10.1080/10978526.2019.1652099}}}\vspace{1mm}

\item{\small Barr\'ia, D., Gonz\'alez-Bustamante, B., \& Cisternas, C. (2019). The Literature on Open Government in Spanish. Analysis of the Dynamics of Production and Citation. {\itshape N\'oesis. Revista de Ciencias Sociales y Humanidades, 28}(56), 22--42. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{http://dx.doi.org/10.20983/noesis.2019.2.3}{\textcolor{blue}{10.20983/noesis.2019.2.3}}}\vspace{1mm}

\item{\small Del Valle, N., \& Gonz\'alez-Bustamante, B. (2018). Political Agenda, Journalism and Digital Media in Chile: Research Notes on Media Pluralism. {\itshape Perspectivas de la Comunicaci\'on, 11}(1), 291--326. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{http://revistas.ufro.cl/ojs/index.php/perspectivas/article/view/1146}{\textcolor{blue}{http://revistas.ufro.cl}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Olivares, A. (2016). Cabinet Turnover and Survival of Ministers in Chile During the Concertacion Governments (1990-2010). {\itshape Colombia Internacional}, 87, 81--108. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{https://doi.org/10.7440/colombiaint87.2016.04}{\textcolor{blue}{10.7440/colombiaint87.2016.04}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. {\itshape et al.} (2016). Civil Service in Chile, Analysis of Senior Civil Servants on the Top Rung of the Hierarchy (2003-13). {\itshape Revista de Administra\c{c}\~ao P\'ublica, 50}(1), 59--79. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{http://dx.doi.org/10.1590/0034-7612145767}{\textcolor{blue}{10.1590/0034-7612145767}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Olivares, A. (2015). Undersecretaries Rotation in Chile: An Exploration of the Second Government Line, 1990-2014. {\itshape Revista de Gesti\'on P\'ublica, IV}(2), 151--190. {\footnotesize \scshape [in Spanish]}. \\ {\scshape doi}: \href{https://doi.org/10.22370/rgp.2015.4.2.2230}{\textcolor{blue}{10.22370/rgp.2015.4.2.2230}}} \vspace{1mm}

\end{benumerate}

\end{publications}

\cvsubsection{ SciELO \& Latindex}

\begin{publications}

\begin{benumerate}{9}
\item{\small Gonz\'alez-Bustamante, B. (2016). Political, Economic, and Intellectual Elites: A Growing Research Agenda for Political Science. {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 54}(1), 7--17. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/article/view/42690}{\textcolor{blue}{https://revistapolitica.uchile.cl}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Cisternas, C. (2016). Political Elites in the Chilean Legislative Branch: The Chamber of Deputies (1990-2014). {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 54}(1), 19--52. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/article/view/42691}{\textcolor{blue}{https://revistapolitica.uchile.cl}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2015). Assessing Twitter as a Public Opinion Measure: An Exploration of the Bachelet's Arrival to the Chilean Presidential Campaign 2013. {\itshape Revista SAAP, 9}(1), 119--141. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{http://ref.scielo.org/dwzhns}{\textcolor{blue}{http://ref.scielo.org}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2015). Electoral Success and Campaign Spending in Elections of Senators and Deputies in Chile 2013. {\itshape Pol\'iticas P\'ublicas, 8}(1), 21--35. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{http://www.revistas.usach.cl/ojs/index.php/politicas/article/view/2182}{\textcolor{blue}{www.revistas.usach.cl}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2014). Popular Election of Regional Councillors 2013. Performance of Political, Family, and Economic Capital in a New Electoral Arena in Chile. {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 52}(2), 49--91. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/article/view/36137}{\textcolor{blue}{https://revistapolitica.uchile.cl}}} \vspace{1mm}

\item{\small Olivares, A. {\itshape et al.} (2014). Think Tanks in the Cabinet: An Exploration of the Chilean Case (2006-2014). {\itshape Revista de Sociolog\'ia,} 29, 37--54. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{https://revistadesociologia.uchile.cl/index.php/RDS/article/view/36177}{\textcolor{blue}{https://revistadesociologia.uchile.cl}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2013). Factors Affecting Access and Permanence of the Governmental Political Elite in Chile (1990-2010). {\itshape Pol\'itica, Revista de Ciencia Pol\'itica, 51}(1), 119--153. {\footnotesize \scshape [in Spanish]}. \\ {\scshape url}: \href{https://revistapolitica.uchile.cl/index.php/RP/article/view/27436}{\textcolor{blue}{https://revistapolitica.uchile.cl}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2013). The Study of the Elites in Chile: Theoretical and Methodological Approaches. {\itshape Intersticios Sociales,} 6, 1--20. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{http://ref.scielo.org/zrnp2k}{\textcolor{blue}{http://ref.scielo.org}}} \vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Henr\'iquez, G. (2012). Digital Campaigns: Branding or Political Participation? {\itshape M\'as Poder Local,} 12, 32--39. {\footnotesize \scshape [in Spanish]}. {\scshape url}: \href{https://www.researchgate.net/publication/260517478_Campanas_digitales_Branding_o_participacion_politica_El_rol_de_las_redes_sociales_en_la_ultima_campana_presidencial_chilena}{\textcolor{blue}{www.researchgate.net}}} \vspace{1mm}

\end{benumerate}

\end{publications}