### 2021-11-25 sections/summary.tex
Update summary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{3mm}
\cvsection{Summary}

\begin{cvparagraph}

I am completing my DPhil (PhD) dissertation in the Department of Politics and International Relations and St Hilda’s College at the University of Oxford, United Kingdom. I am currently based in Oxford, working on my research. Furthermore, I am an Instructor Professor in the Management and Public Policies Department at the Faculty of Administration and Economics at the Universidad de Santiago de Chile.

Before starting the DPhil, in October 2019, I earned an MA (1st) in Political Science and a BA (2:1) in Government, both from the Universidad de Chile. Moreover, I served as a lecturer at the Universidad de Chile and a consultant for the United Nations Development Programme (UNDP).

My doctoral research is supervised by Professor Petra Schleiter and focuses on cabinets’ stability in Latin American presidential democracies. In my DPhil project, I combine the principal-agent theory elements to evaluate the actors’ incentives and strategies in presidential systems. My empirical strategy employs propensity scores and matching methods, competing risks models, and IV estimations. My research is funded by the Chilean National Agency for Research and Development and St Hilda's College.
\vspace{1mm}
\end{cvparagraph}