### 2021-11-25 sections/conferences-full.tex
Update conferences-full.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{List of Conference and Seminar Presentations}

\begin{publications}

\begin{benumerate}{66}

\item{\small González-Bustamante, B. (2022). {\itshape Resignation Calls and Dismissals of Ministers in Latin America}. Presentation delivered at the XXXIX International Congress of the Latin American Studies Association, San Francisco, USA.}\vspace{1mm}

\item{\small González-Bustamante, B. (2022). {\itshape Resignation Calls and Ministerial Turnover in Latin America}. Presentation delivered at the Society for Latin American Studies Annual Conference, Bath, UK.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Cisternas, C. (2022). {\itshape Seniority and Cosponsorship in the Chilean Lower House 2006-2018}. Presentation delivered at the Society for Latin American Studies Annual Conference, Bath, UK.}\vspace{1mm}

\item{\small González-Bustamante, B. (2021). {\itshape The Use of Machine Learning for Public Affairs Research}. Presentation delivered the XII Chilean Congress of Public Administration, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2021). {\itshape Methodological Challenges and Innovations in the Study of Elites}. Presentation delivered at the XLV Annual Conference of the Brazilian National Association of Graduate Studies and Research in Social Sciences, Virtual.}\vspace{1mm}

%% Bamberg

\item{\small González-Bustamante, B. (2021). {\itshape R Applications in Social Sciences: Data Mining and Econometric Modelling}. Presentation delivered at the Political Science Method Lecture at the Universidad de Chile, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2021). {\itshape Vertically Integrated Project: Machine Learning Applications in Public Policy and Economics}. Presentation delivered at the VIP Project Workshop at the Universidad Mayor, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2021). {\itshape Ministerial Stability During Presidential Approval Crises: The Moderating Effect of Ministers' Attributes on Dismissals in Brazil and Chile}. Presentation delivered at the XXVI World Congress of Political Science, Virtual.}\vspace{1mm}

\item{\small Cisternas, C., \& González-Bustamante, B. (2021). {\itshape Political Careers and Co-Sponsorship in the Chilean Lower House 2010-2018}. Presentation delivered at the XXVI World Congress of Political Science, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2021). {\itshape Data Mining, Modelling and Machine Learning Applications for Social Sciences}. Presentation delivered at the I Training Data Lab Workshop, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2020). {\itshape Evolution and Early Governmental Responses to COVID-19 in South America}. Presentation delivered at the Workshop of the School of Political Science and Public Administration at the Universidad de Talca, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2020). {\itshape New Approaches to Studying Political Elites}. Presentation delivered at the Colloquium of the Observatório de Elites at the Universidade Federal do Paraná, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2020). {\itshape Protest and Social Unrest in Latin America: Determinants and Critical Junctures}. Presentation delivered at the Seminar Cycle on Human Rights in Latin America at the Universidade Tiradentes, Virtual.}\vspace{1mm} %% Sergipe, Brazil

\item{\small González-Bustamante, B. (2020). {\itshape Critical Events and Ministerial Turnover in Latin American Presidential Democracies}. Presentation delivered at the PRS DPhil in Politics Workshop in Hilary Term at the University of Oxford, Oxford, UK.}\vspace{1mm}

\item{\small Orchard, X., \& González-Bustamante, B. (2020). {\itshape An Exploration of Profiles and Perceptions in the Diagnosis of Journalistic Practice in Chile}. Presentation delivered at the VI Chilean Association of Communication Researchers Congress, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2019). {\itshape Presidential Stability and Minister Turnover in Argentina, Brazil, and Chile}. Presentation delivered at the XXXVII International Congress of the Latin American Studies Association, Boston, USA.}\vspace{1mm}

\item{\small González-Bustamante, B. (2018). {\itshape Cabinet Stability and Turnover from a Multilevel Approach}. Presentation delivered at the Symposium of Political-Administrative Elites: Planning a Research Agenda at the Universidad de Santiago, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Cisternas, C. (2018). {\itshape Social Network Analysis: Application for Elites Study}. Presentation delivered at the Symposium of Political-Administrative Elites: Planning a Research Agenda at the Universidad de Santiago, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2018). {\itshape Geospatial Analysis of Partisanship Density and Electoral Turnout: Evidence from the Chilean Case}. Presentation delivered at the XIII Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2018). {\itshape Sophistication, Participation, and Political Engagement in Latin America}. Presentation delivered at the VII International Congress of Political Communication, Murcia, Spain.}\vspace{1mm}

\item{\small González-Bustamante, B. (2018). {\itshape Survival of Ministers and Vice-Ministers in Chile: Stability and Turnover from a Multilevel Approach}. Presentation delivered at the International Symposium of Ministers and Ministries: Technocracy and Politics at the Universidade Federal do Paraná, Curitiba, Brazil.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Garrido-Vergara, L. (2018). {\itshape Trajectories and Post-Career of Ministers in Chile}. Presentation delivered at the Research Seminar of the Institute of Public Affairs at the Universidad de Chile, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Del Valle, N. (2017). {\itshape Media Pluralism and Digital Media Public Opinion}. Presentation delivered at the III Seminar of Pluralism and Mass Media in Chile at the Universidad Alberto Hurtado, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., Maillet, A., \& Olivares, A. (2017). {\itshape Revolving Door? Analysis of Public-Private Circulation in Chile}. Presentation delivered at the II Foreign Affairs Meeting at the Universidad de Santiago, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Garrido-Vergara, L. (2017). {\itshape Socialisation Spaces and Ministerial Post-Career in the Chilean Transition and Post-Transition}. Presentation delivered at the IX Chilean Congress of Public Administration, Concepcion, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., Maillet, A., \& Olivares, A. (2017). {\itshape Public-Private: An Especially Tenuous Frontier. Political-Professional Trajectories of the Chilean Authorities (2000-2014)}. Presentation delivered at the IX Latin American Congress of Political Science, Montevideo, Uruguay.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Joignant, A. (2016). {\itshape Gaps and Elite-Mass Congruence in Chile and Uruguay}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities at the Universidad de Santiago, Santiago, Chile.}\vspace{1mm}

\item{\small Joignant, A., \& González-Bustamante, B. (2016). {\itshape Family Ties in the Senate and the Chamber of Deputies, Chile from 1989 to 2014}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities at the Universidad de Santiago, Santiago, Chile.}\vspace{1mm}

\item{\small Barría, D., González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Analysis of the Academic Production on Open Government in Spanish}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities at the Universidad de Santiago, Santiago, Chile.}\vspace{1mm}

\item{\small Maillet, A., González-Bustamante, B., \& Olivares, A. (2016). {\itshape Revolving Door and Public-Private Circulation in Chile}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities at the Universidad de Santiago, Santiago, Chile.}\vspace{1mm}

\item{\small Barría, D., González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Open Government in Latin America and Spain: Analysis of the Scientific Production in Spanish}. Presentation delivered at the XXI CLAD International Congress on the State Reform and Public Administration, Santiago, Chile.}\vspace{1mm}

\item{\small Joignant, A., \& González-Bustamante, B. (2016). {\itshape Performance of Family Political Capital in the Chilean Post-Authoritarianism}. Presentation delivered at the XII Chilean Congress of Political Science, Pucon, Chile.}\vspace{1mm}

\item{\small Olivares, A., \& González-Bustamante, B. (2016). {\itshape Turnover and Ministerial Survival Before the Chilean Democratic Breakdown}. Presentation delivered at the XII Chilean Congress of Political Science, Pucon, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Barría, D. (2016). {\itshape Production Patterns of the Contemporary Knowledge in Social Sciences}. Presentation delivered at the IX Chilean Congress of Sociology, Talca, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Co-Authorship Networks of the Political Science in Spanish (2010-2015)}. Presentation delivered at the IX Chilean Congress of Sociology, Talca, Chile.}\vspace{1mm}

\item{\small Cisternas, C., González-Bustamante, B., \& Barría, D. (2016). {\itshape Analysis of the Scientific Production in Spanish on Public Administration and Public Policies (2010-2012)}. Presentation delivered at the IX Chilean Congress of Sociology, Talca, Chile.}\vspace{1mm}

\item{\small Cisternas, C., \& González-Bustamante, B. (2016). {\itshape The CIEPLAN Monks: Intellectual Production and Co-Citation Networks During the Eighties}. Presentation delivered at the International Seminar of Social Sciences at the Crossroad: Latin American Intellectuals and Technocrats in Authoritarianism Times at the Universidad Bernardo O'Higgins, Santiago, Chile.}\vspace{1mm}

\item{\small Alenda, S., \& González-Bustamante, B. (2016). {\itshape The Reshaping of the Political Right in Chile: Change and Continuity in the Perceptions of Social Inequalities}. Presentation delivered at the III ISA Forum of Sociology, Vienna, Austria.}\vspace{1mm}

\item{\small González-Bustamante, B. (2016). {\itshape Representation and Elite-Mass Congruence in Chile and Uruguay}. Presentation delivered at the XXXIV International Congress of the Latin American Studies Association, New York, USA.}\vspace{1mm}

\item{\small Alenda, S., González-Bustamante, B., \& Gartenlaub, A. (2016). {\itshape The Reshaping of the Political Right in Chile}. Presentation delivered at the XXXIV International Congress of the Latin American Studies Association, New York, USA.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Barría, D. (2015). {\itshape Stateness in the Chilean Post-Transition}. Presentation delivered at the VIII Chilean Congress of Public Administration, Rancagua, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2015). {\itshape Media Consumption and Political Participation in Argentina, Chile, and Uruguay}. Presentation delivered at the XII Argentinian Congress of Political Science, Mendoza, Argentina.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Olivares, A. (2015). {\itshape Trajectories of Ministers and Undersecretaries in Chile During Concertacion Governments (1990-2010)}. Presentation delivered at the VIII Latin American Congress of Political Science, Lima, Peru.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Barría, D. (2015). {\itshape Stateness' Support in Chile from 2007 to 2014}. Presentation delivered at the VIII Latin American Congress of Political Science, Lima, Peru.}\vspace{1mm}

\item{\small Olivares, A., \& González-Bustamante, B. (2015). {\itshape Survival of Ministers and Undersecretaries in Chile (1990-2014)}. Presentation delivered at the International Congress of Elites and Leadership at the Universidad  de Salamanca, Salamanca, Spain.}\vspace{1mm}

\item{\small Garrido-Vergara, L., \& González-Bustamante, B. (2015). {\itshape Elites and Political Reforms: The Chilean Case}. Presentation delivered at the International Congress of Elites and Leadership at the Universidad  de Salamanca, Salamanca, Spain.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Ruiz, I. (2014). {\itshape Radiography to the Popular Election of Regional Councillors 2013}. Presentation delivered at the XI Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{\small Thezá, M., \& González-Bustamante, B. (2014). {\itshape Dimensions of the Young People Multi-Associative Participation in Chile}. Presentation delivered at the XI Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2014). {\itshape Social Media and Malaise with Democracy in Argentina and Uruguay}. Presentation delivered at the VI Latin American Congress of Public Opinion, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2014). {\itshape Media Consumption and Political Participation in the Southern Cone}. Presentation delivered at the VI Latin American Congress of Public Opinion, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2014). {\itshape Chilean Elections 2013: Incumbency, Campaign Spending, and Electoral Success in the Congress}. Presentation delivered at the Centre for Political Analysis and Research Colloquium, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2013). {\itshape Digital Independent Media in Chile}. Presentation delivered at the Seminar of Mass Media Pluralism in Chile at the Centre for Political Analysis and Research, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2013). {\itshape The Governmental Political Elite in Chile}. Presentation delivered at the Seminar of Elites Studies in Chile at the Centre for Political Analysis and Research, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2013). {\itshape Regional Barometer 2013: Regional Public Opinion in Bio-Bio, Los Lagos, and Los Rios}. Presentation delivered at the VI Chilean Congress of Public Administration, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2013). {\itshape Online Social Networks and Political Behaviour in Chile: Twitter as a Measure of Public Opinion During the Primary Presidential Elections in 2013}. Presentation delivered at the II International Congress of Political Communication, Toluca, Mexico.}\vspace{1mm}

\item{\small Chávez, D., \& González-Bustamante, B. (2013). {\itshape Partisanship Identification and Disaffection in Ecuador}. Presentation delivered at the XXIX Latin American Congress of Sociology, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Soto, D. (2013). {\itshape Digital Independent Media in Chile: Democratisation in the Media System?} Presentation delivered at the VII Latin American Congress of Political Science, Bogota, Colombia.}\vspace{1mm}

\item{\small González-Bustamante, B. (2013). {\itshape Trajectories and Patterns of the Ministers' Political Careers in Chile}. Presentation delivered at the XI Argentinian Congress of Political Science, Parana, Argentina.}\vspace{1mm}

\item{\small González-Bustamante, B. (2013). {\itshape Twitter as Public Opinion Measure: The Bachelet's Arrival to the Chilean Presidential Campaign 2013}. Presentation delivered at the XI Argentinian Congress of Political Science, Parana, Argentina.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Valenzuela, P. (2012). {\itshape Representation Gaps and Incorporated Capital: The Case of the Concertacion's Political Youths}. Presentation delivered at the VII Chilean Congress of Sociology, Pucon, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2012) {\itshape The Transversal Political Party Thesis: Beyond the Myth}. Presentation delivered at the X Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Chávez, D. (2012). {\itshape Bolivian Inter-Cultural Democracy: Advances and Challenges}. Presentation delivered at the X Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{\small González-Bustamante, B. (2012). {\itshape Lights and Shadows of the Public Management Professionalisation in the Chilean Case}. Presentation delivered at the XXII World Congress of Political Science, Madrid, Spain.}\vspace{1mm}

\item{\small Ríos, T. E., \& González-Bustamante, B. (2012). {\itshape Media System and Democratisation in the Chilean Case}. Presentation delivered at the XXII World Congress of Political Science, Madrid, Spain.}\vspace{1mm}

\item{\small González-Bustamante, B., \& Henríquez, G. (2012). {\itshape The Role of the Online Social Networks in the Last Chilean Presidential Campaign}. Presentation delivered at the I International Congress of Political Communication, Madrid, Spain.}\vspace{1mm}

\item{\small González-Bustamante, B. (2012). {\itshape Research Notes and Reflections on the Study of Elites in Chile}. Presentation delivered at the VI Latin American Congress of Political Science, Quito, Ecuador.}

\end{benumerate}

\end{publications}