### 2022-03-10 spanish/sections/reviewer.tex
Update reviewer.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in Spanish
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Revisiones como árbitro \& experto}

\begin{publications}

\begin{itemize}
\item{\small Proyecto V-Dem, Universidad de Gotemburgo, Suecia, 2014, 2018-2020}
\item{\small Encuesta de Reputación Académica Global, Clarivate Analytics, EUA, 2020}
\item{\small Comisión Sectorial de Investigación, Universidad de la Rep\'ublica, Uruguay, 2016, 2018}
\item{\small Times Higher Education Ranking, {\scshape Elsevier}, Países Bajos, 2018-2019, 2022}
\item{\small Revisor en revistas {\scshape WoS-ISI/SSCI/ESCI} \& {\scshape Scopus}, desde 2013}
\item{\small Revisor en revistas {\scshape SciELO} \& {\scshape Latindex}, desde 2013}
\end{itemize}

\vspace{1mm}
\end{publications}