### 2022-04-08 sections/working-papers.tex
Update working-papers.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% Creative Commons Attribution 4.0 International License
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Data Collection Projects}

\begin{publications}

\begin{benumerate}{2}

\item{\small Gonz\'alez-Bustamante, B., \& Olivares, A. (2022). Data Set on Chilean Ministers (1990-2014) (Version 3.3.6 -- Dry Bonus) [Data set]. {\scshape doi:} \href{https://doi.org/10.5281/zenodo.5744536}{\textcolor{blue}{10.5281/zenodo.5744536}}}\vspace{1mm}

\item{Gonz\'alez-Bustamante, B., \& Olivares, A. (2022). Data Set on Chilean Undersecretaries (1990-2022) (Version 2.1.3 -- Shiny Snow) [Data set]. {\scshape doi:} \href{https://doi.org/10.5281/zenodo.5715384}{\textcolor{blue}{10.5281/zenodo.5715384}}}\vspace{1mm}

\end{benumerate}

\end{publications}

\cvsection{Working Papers and Reviews}

\begin{publications}

\begin{benumerate}{4}

\item{\small Maillet, A., Gonz\'alez-Bustamante, B., \& Olivares, A. (2016). Revolving Door. Analysis of Public-Private Circulation in Chile (2000-2014). {\itshape Serie Documentos de Trabajo PNUD-Desigualdad}, 7, 1--40. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{http://doi.org/10.13140/RG.2.2.25510.42566}{\textcolor{blue}{10.13140/RG.2.2.25510.42566}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Olivares, A. (2016). Undersecretaries Rotation in the Chilean Cabinets (1990-2014). {\itshape Observat\'orio de elites pol\'iticas e sociais do Brasil. NUSP/UFPR, 3}(5), 1--19. {\footnotesize \scshape [in Portuguese]}. {\scshape url}: \href{https://www.researchgate.net/publication/321993740_A_rotatividade_de_subsecretarios_dos_ministerios_no_Chile_1990-2014}{\textcolor{blue}{www.researchgate.net}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B., \& Soto, D. (2016). Pluralism in the Chilean Digital Mass Media, Exploratory Study of Key Cases. {\itshape Serie Pol\'itica, Gesti\'on y Pol\'iticas P\'ublicas}, 2, 1--26. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{http://doi.org/10.13140/RG.2.2.33899.03362/1}{\textcolor{blue}{10.13140/RG.2.2.33899.03362/1}}}\vspace{1mm}

\item{\small Gonz\'alez-Bustamante, B. (2015). Programas de Naciones Unidas para el Desarrollo (PNUD). (2014). Ciudadanía Política: Voz y Participación Ciudadana en América Latina. Buenos Aires: Siglo Veintiuno Editores. {\itshape Revista de Gestión Pública, IV}(1), 143--149. {\footnotesize \scshape [in Spanish]}. {\scshape doi}: \href{https://doi.org/10.22370/rgp.2015.4.1.2240}{\textcolor{blue}{10.22370/rgp.2015.4.1.2240}}}\vspace{1mm}

\end{benumerate}

\end{publications}