### 2022-09-01 sections/summary.tex
Update summary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{3mm}
\cvsection{Summary}

\begin{cvparagraph}

I am completing my DPhil (PhD) dissertation in the Department of Politics and International Relations and St Hilda’s College at the University of Oxford, United Kingdom. Before starting the DPhil, in October 2019, I earned an MA (1st) in Political Science and a BA (2:1) in Government, both from the Universidad de Chile. Moreover, I served as a lecturer at the Universidad de Santiago and Universidad de Chile and as a consultant for the United Nations Development Programme (UNDP).

My doctoral research is supervised by Professor Petra Schleiter and focuses on cabinets’ stability in Latin American presidential democracies. In my DPhil project, I combine the principal-agent theory elements to evaluate the actors’ incentives and strategies in presidential systems. My data gathering strategy is based on archives revision using Optical Character Recognition (OCR), Natural Language Processing (NLP), and machine learning techniques. In addition, my empirical strategy employs propensity scores and matching methods, competing risks models, and IV estimations.

My research is funded by the Chilean National Agency for Research and Development, the St Hilda’s College Muriel Wise Fund and the Society for Latin American Studies Postgraduate and Postdoctoral Research Award.
\vspace{1mm}
\end{cvparagraph}