### 2022-11-13 sections/oxf_refs.tex
Create oxf_refs.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Referees}

\begin{publications}

\begin{itemize}
\item[]{\small {\bfseries Petra Schleiter}, Professor of Comparative Politics, University of Oxford, UK}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:petra.schleiter@politics.ox.ac.uk}{\textcolor{blue}{petra.schleiter@politics.ox.ac.uk}}}} \vspace{1.5mm}
\item[]{\small {\bfseries David Doyle}, Professor of Politics, University of Oxford, UK}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:david.doyle@politics.ox.ac.uk}{\textcolor{blue}{david.doyle@politics.ox.ac.uk}}}} \vspace{1.5mm}
\end{itemize}

\end{publications}