### 2023-03-04 short/sections/selected-papers.tex
Create selected-papers.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Featured Publications}

\begin{publications}

\begin{itemize}

\item{González-Bustamante, B. (2022). Ministerial Stability During Presidential Approval Crises: The Moderating Effect of Ministers' Attributes on Dismissals in Brazil and Chile. {\itshape The British Journal of Politics and International Relations}. OnlineFirst. {\scshape doi:} \href{https://doi.org/10.1177/13691481221124850}{\textcolor{blue}{10.1177/13691481221124850}}.}\vspace{1mm}

\item{Fleming, T. G., González-Bustamante, B., \& Schleiter, P. (2022). Cabinet Reshuffles and Parliamentary No-Confidence Motions. {\itshape Government and Opposition}. FirstView. {\scshape doi:} \href{https://doi.org/10.1017/gov.2022.23}{\textcolor{blue}{10.1017/gov.2022.23}}.}\vspace{1mm}

\item{Orchard, X., \& González-Bustamante, B. (2022). Power Hierarchies and Visibility in the News: Exploring Determinants of Politicians’ Presence and Prominence in the Chilean Press (1991-2019). {\itshape The International Journal of Press/Politics}. OnlineFirst. {\scshape doi:} \href{https://doi.org/10.1177/19401612221089482}{\textcolor{blue}{10.1177/19401612221089482}}.}\vspace{1mm}

\item{\small González-Bustamante, B. (2021). Evolution and Early Government Responses to COVID-19 in South America. {\itshape World Development, 137}, 105180. {\scshape doi:} \href{https://doi.org/10.1016/j.worlddev.2020.105180}{\textcolor{blue}{10.1016/j.worlddev.2020.105180}}.}\vspace{1mm}

\item{\small González-Bustamante, B., Astete, M., \& Orvenes, B. (2020). Senior Public Managers: A Novel Dataset on Members of the Chilean Civil Service. {\itshape Revista de Gestión Pública, IX}(2), 151--169. {\footnotesize \scshape [in Spanish and English]}. {\scshape doi:} \href{https://doi.org/10.22370/rgp.2020.9.2.2920}{\textcolor{blue}{10.22370/rgp.2020.9.2.2920}}. {\scshape \footnotesize SocArXiv:} \href{https://doi.org/10.31235/osf.io/vshcz}{\textcolor{blue}{10.31235/osf.io/vshcz}}.}\vspace{1mm}

\item{\small González-Bustamante, B. (2020). The Politics-Administration Dichotomy: A Case Study of the Chilean Executive During the Democratic Post-Transition. {\itshape Bulletin of Latin American Research, 39}(5), 582--597. \\ {\scshape doi:} \href{https://doi.org/10.1111/blar.13044}{\textcolor{blue}{10.1111/blar.13044}}.}\vspace{1mm}

\item{\small González-Bustamante, B., Carvajal, A., \& González, A. (2020). Determinants of E-Government in the Municipalities: Evidence from the Chilean Case. {\itshape Gestión y Política Pública, XXIX}(1), 97--129. {\footnotesize \scshape [in Spanish]}. \\ {\scshape doi:} \href{http://dx.doi.org/10.29265/gypp.v29i1.658}{\textcolor{blue}{10.29265/gypp.v29i1.658}}. {\scshape \footnotesize SocArXiv:} \href{https://doi.org/10.31235/osf.io/fze3x}{\textcolor{blue}{10.31235/osf.io/fze3x}}.} \vspace{1mm}

\item{\small Maillet, A., González-Bustamante, B., \& Olivares, A. (2019). Public-Private Circulation and the Revolving Door in the Chilean Executive Branch (2000-2014). {\itshape Latin American Business Review, 20}(4), 367--387. \\ {\scshape doi:} \href{https://doi.org/10.1080/10978526.2019.1652099}{\textcolor{blue}{10.1080/10978526.2019.1652099}}.}\vspace{1mm}

\item{\small González-Bustamante, B. (2019). Gaps, Representation, and Elite-Mass Congruence in Chile and Uruguay. {\itshape Convergencia. Revista de Ciencias Sociales}, 80, 1--27. {\footnotesize \scshape [in Spanish]}. {\scshape doi:} \href{https://doi.org/10.29101/crcs.v26i80.11097}{\textcolor{blue}{10.29101/crcs.v26i80.11097}}. {\scshape \footnotesize SocArXiv:} \href{https://doi.org/10.31235/osf.io/cqym8}{\textcolor{blue}{10.31235/osf.io/cqym8}}.}\vspace{1mm} %% {\scshape url:}  \href{https://www.redalyc.org/jatsRepo/105/10559568002/index.html}{\textcolor{blue}{www.redalyc.org}}

\item{\small González-Bustamante, B. (2018). Civil Service Models in Latin America. In A. Farazmand (ed.), {\itshape Global Encyclopedia of Public Administration, Public Policy, and Governance}. Cham: Springer. {\scshape doi:} \href{https://doi.org/10.1007/978-3-319-20928-9\_2699}{\textcolor{blue}{10.1007/978-3-319-20928-9\_2699}}. {\scshape \footnotesize SocArXiv:} \href{https://doi.org/10.31235/osf.io/mp4qd}{\textcolor{blue}{10.31235/osf.io/mp4qd}}.}\vspace{1mm}

\end{itemize}

\end{publications}
%% \pagebreak