### 2023-09-22 sections/languages.tex
Update languages.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Languages \& Software}

\begin{cvskills}

\cvskill
{Languages} 
{Spanish (native speaker), English (fluent), Portuguese (advanced reading skills), Italian}

\cvskill
{} 
{(elemental reading skills), French (elemental reading skills)} \\

\cvskill
{Programming \&} 
{R, Python, Stata, {\scshape spss/pspp}, Gephi, {\scshape ucinet}, NetDraw, Tosmana, Fuzzy-Sets, ArcGIS, {\scshape atlas}.ti,}

\cvskill
{Software} 
{EndNote, Zotero, {\scshape HPoP}, {\LaTeX}, Bib{\TeX}, JabRef, MS Office, WordPress, Moodle, {\scshape html}, {\scshape css}, {\scshape xhtml}} \\

\end{cvskills}