### 2023-09-23 sections/selected-training.tex
Update selected-training.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Selected Certifications \& Training}

%% OX|BER
%% USACH-Nuffield
%% R Conference

\begin{cvhonors}
\cvconf
{Research Data Workshop IT Centre at the University of Oxford} 
{Oxford, GBR}
{2019}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Longitudinal Networks with RSiena Workshop at the P Universidad Cat\'olica}
{Santiago, CHL}
{2018}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Universidad de Santiago Winter School for Academics}
{Santiago, CHL}
{2016}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Nuffield-USACH CEES Experimental Summer School}
{Santiago, CHL}
{2016}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Universidad de Santiago Winter School for Academics}
{Santiago, CHL}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Latin American Political Science Association Methodological School}
{Lima, PER}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{Universidad Católica de Uruguay Winter School in Methods and Data Analysis}
{Montevideo, URY}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvconf
{P Universidad Caólica Summer School in Mixed Methods}
{Santiago, CHL}
{2014}
\end{cvhonors}

\vspace{1mm}