### 2023-09-23 spanish/sections/deprecated/training.tex
Create training.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Certification \& Training}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{DPhil Traning Programme Trinity Term at the University of Oxford} 
{Training in geospatial modelling} 
{Oxford, GBR}
{2020}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{DPhil Traning Programme Hilary Term at the University of Oxford} 
{Training in formal analysis and game theory, randomised trials, regression and matching, instrumental variables, regression discontinuity, differences-in-differences, and placebo tests} 
{Oxford, GBR}
{2020}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Research Data Workshop IT Centre at the University of Oxford} 
{Training in managing research data} 
{Oxford, GBR}
{2019}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{DPhil Traning Programme Michaelmas Term at the University of Oxford}
{Training in research design, logit and OLS models, multinomial models, ordinal logistic regressions,  OLS regularisation and cross-validation procedures, shrinkage methods, regression tree, random forest algorithm, principal component analysis based on unsupervised learning, and count models}
{Oxford, GBR}
{2019}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Longitudinal Networks with RSiena Workshop at the Pontifica Universidad Cat\'olica}
{Training in stochastic actor oriented models in R}
{Santiago, CHL}
{2018}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Universidad de Santiago Winter School for Academics}
{Training in evaluation techniques and teaching strategies for academic writing}
{Santiago, CHL}
{2016}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Nuffield-USACH CEES Experimental Summer School}
{Training in experiments in social sciences}
{Santiago, CHL}
{2016}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Universidad de Santiago Winter School for Academics}
{Training in qualitative data analysis in {\scshape atlas}.ti}
{Santiago, CHL}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Latin American Political Science Association Methodological School}
{Training in programming in Python, social network analysis in R and Python, and qualitative comparative analysis in R}
{Lima, PER}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Universidad Cat\'olica de Uruguay Winter School in Methods and Data Analysis}
{Training in maths for social sciences, programming in R and Stata, Bayesian statistics, and logit and probit models in R}
{Montevideo, URY}
{2015}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Pontificia Universidad Ca\'olica Summer School in Mixed Methods}
{Training in set theory and logic, case studies and multi-methods, process tracing, and experiments in social sciences}
{Santiago, CHL}
{2014}
\end{cvhonors}

\vspace{1mm}