### 2023-09-23 sections/conferences-full.tex
Update conferences-full.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{List of Conference and Seminar Presentations}

\begin{publications}

\begin{benumerate}{87}

\item{González-Bustamante, B. (2023). {\itshape Corrective Effect of Ministerial Terminations on Presidential Approval}. Presentation delivered at the ECPR General Conference (ECPR), Charles University, Prague, Czech Republic.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Data Mining and Machine Learning for Political Events: An Application to Produce Indicators for Presidential Systems}. Presentation delivered at the ECPR General Conference (ECPR), Charles University, Prague, Czech Republic.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape The Impact of Ministers' Profiles on Cabinet Turnover during Presidential Crises}. Presentation delivered at the XXVII World Congress of Political Science (IPSA), Buenos Aires, Argentina.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Corrective Effect of Ministerial Terminations on Presidential Approval}. Presentation delivered at the Political Studies Association Early Career Network Annual Conference ``Political Worlds'' (PSA-ECN), Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Propensity score matching para datos observacionales}. Presentation delivered at the Doctorate in Social Complexity Sciences, Universidad del Desarrollo, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Quantitative Methods for Studying Elites: Ministers' Profiles and Cabinet Turnover in Presidential Crises}. Presentation delivered at the Workshop Elites and Class -- Interrogating the Rich and Powerful in Latin America and Beyond, University of Kassel, Kassel, Alemania.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Técnicas de propensión y algoritmos de emparejamiento para datos observacionales}. Presentation delivered at the Fundación Analítica's Monthly Workshops, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Resignation Calls and Ministerial Turnover in Presidential Systems}. Presentation delivered at the European Consortium for Political Research Joint Sessions of Workshops (ECPR), Institut d'Études Politiques de Toulouse, Toulouse, France.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Resignation Calls, Reallocations and Individual Ministerial Terminations in Latin America}. Presentation delivered at the 73rd Political Studies Association Annual International Conference (PSA), Liverpool, United Kingdom.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Resignation Calls, Reallocations and Individual Ministerial Terminations in Latin America}. Presentation delivered at the Society for Latin American Studies Annual Conference (SLAS), Belfast, United Kingdom.}\vspace{1mm}

%% \item{González-Bustamante, B. (2022). {\itshape Machine Learning and Political Events: Application of Semisupervised Approach to Produce a Data Set on Latin American Cabinets}. Presentation delivered at the II Training Data Lab Workshop, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2022). {\itshape Aprendizaje automático y eventos políticos: Aplicación de un enfoque semisupervisado para producir un conjunto de datos sobre gabinetes latinoamericanos}. Presentation delivered at the II Training Data Lab Workshop, Virtual.}\vspace{1mm}

\item{González-Bustamante, B., \& Aguilar, D. (2022). {\itshape Digital Development and Open Government: Evidence from Chilean Local Governments}. Presentation delivered at the Digital Democracy Workshop, Digital Society Initiative, University of Zurich, Zurich, Switzerland.}\vspace{1mm}

%% \item{González-Bustamante, B. (2022). {\itshape Machine Learning for the Study of Political Events: Application of Semisupervised Models on Latin American Archives}. Presentation delivered at the LatinR Conference, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2022). {\itshape Aprendizaje automático para el estudio de eventos políticos: Aplicación de modelos semisupervisados en archivos de América Latina}. Presentation delivered at the LatinR Conference, Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Aguilar, D. (2022). {\itshape Integrated Repository on Local Government Indicators in Chile}. Presentation delivered at the LatinR Conference, Virtual.}\vspace{1mm}

\item{González-Bustamante, B., \& Aguilar, D. (2022). {\itshape Repositorio integrado de indicadores de gobiernos locales en Chile}. Presentation delivered at the LatinR Conference, Virtual.}\vspace{1mm}

\item{González-Bustamante, B., \& Aguilar, D. (2022). {\itshape Territorial Patterns of Open E-Government: Evidence from Chilean Municipalities}. Presentation delivered at the International Political Science Association Seminar Smart Cities and Open Government: Local Responses to Participation, Public Information, and Democracy, Madrid, Spain}\vspace{1mm}

%% \item{González-Bustamante, B. (2022). {\itshape Quantitative Methods for Studying Elites: Practical Applications, Biases and Potential}. Presentation delivered at the XI Latin American Congress of Political Science, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2022). {\itshape Métodos cuantitativos para estudiar a las élites: Aplicaciones prácticas, sesgos y potencialidades}. Presentation delivered at the XI Latin American Congress of Political Science (ALACIP), Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Aguilar, D. (2022). {\itshape Territorial Patterns in Digital Development and Open Government: Evidence from Chilean Municipalities}. Presentation delivered at the XI Latin American Congress of Political Science, Virtual.}\vspace{1mm}

\item{González-Bustamante, B., \& Aguilar, D. (2022). {\itshape Patrones territoriales en el desarrollo digital y el gobierno abierto: Evidencia de los municipios chilenos}. Presentation delivered at the XI Latin American Congress of Political Science (ALACIP), Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2022). {\itshape Ministerial Stability During Presidential Approval Crises: Evidence from Brazil and Chile}. Presentation delivered at the Political Studies Association Early Career Network Annual Conference ``The Time is Now'' (PSA-ECN), Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2022). {\itshape Time-Dependent Data Encoding and Time-Varying Exposure Survival Models to Study Presidential Crises}. Presentation delivered at the R User Annual Conference, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2022). {\itshape Resignation Calls and Dismissals of Ministers in Latin America: Data Gathering using NLP and Machine Learning}. Presentation delivered at the XXXIX International Congress of the Latin American Studies Association (LASA), Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2022). {\itshape Insights of a Number of Ongoing Research on Elites in Latin America: Topics, Methods and Possibilities for Comparison}. Presentation delivered at the XXXIX International Congress of the Latin American Studies Association (LASA), Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2022). {\itshape Resignation Calls and Ministerial Turnover in Latin America: Notes on Data Gathering using Machine Learning}. Presentation delivered at the Society for Latin American Studies Annual Conference (SLAS), Bath, United Kingdom.}\vspace{1mm}

\item{González-Bustamante, B., \& Cisternas, C. (2022). {\itshape Seniority and Cosponsorship in the Chilean Lower House 2006-2018}. Presentation delivered at the Society for Latin American Studies Annual Conference (SLAS), Bath, United Kingdom.}\vspace{1mm}

%% \item{González-Bustamante, B. (2021). {\itshape The Use of Machine Learning for Public Affairs Research}. Presentation delivered the XII Chilean Congress of Public Administration, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2021). {\itshape El uso de machine learning para la investigación en asuntos públicos}. Presentation delivered the XII Chilean Congress of Public Administration, Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B. (2021). {\itshape Methodological Challenges and Innovations in the Study of Elites}. Presentation delivered at the XLV Annual Conference of the Brazilian National Association of Graduate Studies and Research in Social Sciences, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2021). {\itshape Desafíos metodológicos e innovaciones en el estudio de las élites}. Presentation delivered at the XLV Annual Conference of the Brazilian National Association of Graduate Studies and Research in Social Sciences (ANPOCS), Virtual.}\vspace{1mm}

\item{Fleming, T. G., González-Bustamante, B., \& Schleiter, P. (2021). {\itshape Cabinet Reshuffles and Parliamentary No-Confidence Motions}. Presentation delivered at the Workshop on Coalition Politics and Portfolio Design, University of Bamberg, Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B. (2021). {\itshape R Applications in Social Sciences: Data Mining and Econometric Modelling}. Presentation delivered at the Political Science Method Lecture, Universidad de Chile, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2021). {\itshape Aplicaciones de R en ciencias sociales: Minería de datos y modelamiento econométrico}. Presentation delivered at the Political Science Method Lecture, Universidad de Chile, Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B. (2021). {\itshape Vertically Integrated Project: Machine Learning Applications in Public Policy and Economics}. Presentation delivered at the VIP Project Workshop, Universidad Mayor, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2021). {\itshape Aplicaciones de Machine Learning en políticas públicas y economía}. Presentation delivered at the VIP Project Workshop, Universidad Mayor, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2021). {\itshape Ministerial Stability During Presidential Approval Crises: The Moderating Effect of Ministers' Attributes on Dismissals in Brazil and Chile}. Presentation delivered at the XXVI World Congress of Political Science (IPSA), Virtual.}\vspace{1mm}

\item{Cisternas, C., \& González-Bustamante, B. (2021). {\itshape Political Careers and Co-Sponsorship in the Chilean Lower House 2010-2018}. Presentation delivered at the XXVI World Congress of Political Science (IPSA), Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B. (2021). {\itshape Data Mining, Modelling and Machine Learning Applications for Social Sciences}. Presentation delivered at the I Training Data Lab Workshop, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2021). {\itshape Aplicaciones de data mining, modelamiento y machine learning para ciencias sociales}. Presentation delivered at the I Training Data Lab Workshop, Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B. (2020). {\itshape Evolution and Early Governmental Responses to COVID-19 in South America}. Presentation delivered at the Workshop of the School of Political Science and Public Administration, Universidad de Talca, Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2020). {\itshape Evolución y respuestas gubernamentales tempranas al COVID-19 en América del Sur}. Presentation delivered at the Workshop of the School of Political Science and Public Administration, Universidad de Talca, Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B. (2020). {\itshape New Approaches to Studying Political Elites}. Presentation delivered at the Colloquium of the Observatório de Elites, Universidade Federal do Paraná, Virtual.}\vspace{1mm}

\item{\small González-Bustamante, B. (2020). {\itshape Nuevos enfoques para estudiar élites políticas}. Presentation delivered at the Colloquium of the Observatório de Elites, Universidade Federal do Paraná, Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B. (2020). {\itshape Protest and Social Unrest in Latin America: Determinants and Critical Junctures}. Presentation delivered at the Seminar Cycle on Human Rights in Latin America, Universidade Tiradentes, Virtual.}\vspace{1mm} %% Sergipe, Brazil

\item{González-Bustamante, B. (2020). {\itshape Protesta y malestar social en América Latina: Determinantes y coyunturas críticas}. Presentation delivered at the Seminar Cycle on Human Rights in Latin America, Universidade Tiradentes, Virtual.}\vspace{1mm}

%% \item{González-Bustamante, B. (2020). {\itshape Critical Events and Ministerial Turnover in Latin American Presidential Democracies}. Presentation delivered at the Pembroke College PRS DPhil in Politics Workshop, University of Oxford, Oxford, UK.}\vspace{1mm}

\item{González-Bustamante, B. (2020). {\itshape Critical Events and Ministerial Turnover in Latin American Presidential Democracies}. Presentation delivered at the Pembroke College Workshop, University of Oxford, Oxford, United Kingdom.}\vspace{1mm}

%% \item{Orchard, X., \& González-Bustamante, B. (2020). {\itshape An Exploration of Profiles and Perceptions in the Diagnosis of Journalistic Practice in Chile}. Presentation delivered at the VI Chilean Association of Communication Researchers Congress, Virtual.}\vspace{1mm}

\item{Orchard, X., \& González-Bustamante, B. (2020). {\itshape Una exploración de perfiles y percepciones en el diagnóstico sobre la práctica periodística en Chile}. Presentation delivered at the VI Chilean Association of Communication Researchers Congress (INCOM), Virtual.}\vspace{1mm}

\item{González-Bustamante, B. (2019). {\itshape Presidential Stability and Minister Turnover in Argentina, Brazil, and Chile}. Presentation delivered at the XXXVII International Congress of the Latin American Studies Association (LASA), Boston, United States.}\vspace{1mm}

%% \item{González-Bustamante, B. (2018). {\itshape Cabinet Stability and Turnover from a Multilevel Approach}. Presentation delivered at the Symposium of Political-Administrative Elites: Planning a Research Agenda, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2018). {\itshape Supervivencia de ministros y viceministros en Chile: Estabilidad y rotación en una lógica multinivel}. Presentation delivered at the Symposium of Political-Administrative Elites: Planning a Research Agenda, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Cisternas, C. (2018). {\itshape Social Network Analysis: Application for Elites Study}. Presentation delivered at the Symposium of Political-Administrative Elites: Planning a Research Agenda, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Cisternas, C. (2018). {\itshape ForceAtlas2, un algoritmo de diseño gráfico continuo: Aplicación para el estudio de las élites}. Presentation delivered at the Symposium of Political-Administrative Elites: Planning a Research Agenda, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2018). {\itshape Geospatial Analysis of Partisanship Density and Electoral Turnout: Evidence from the Chilean Case}. Presentation delivered at the XIII Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2018). {\itshape Análisis geoespacial de la densidad partidista y la participación electoral: Evidencia del caso chileno}. Presentation delivered at the XIII Chilean Congress of Political Science (ACCP), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2018). {\itshape Sophistication, Participation, and Political Engagement in Latin America}. Presentation delivered at the VII International Congress of Political Communication, Murcia, Spain.}\vspace{1mm}

\item{González-Bustamante, B. (2018). {\itshape Sofisticación, participación y compromiso político en América Latina}. Presentation delivered at the VII International Congress of Political Communication (ALICE), Murcia, Spain.}\vspace{1mm}

%% \item{González-Bustamante, B. (2018). {\itshape Survival of Ministers and Vice-Ministers in Chile: Stability and Turnover from a Multilevel Approach}. Presentation delivered at the International Symposium of Ministers and Ministries: Technocracy and Politics at the Universidade Federal do Paraná, Curitiba, Brazil.}\vspace{1mm}

\item{González-Bustamante, B. (2018). {\itshape Supervivencia de ministros y viceministros en Chile: Estabilidad y rotación en una lógica multinivel}. Presentation delivered at the International Symposium of Ministers and Ministries: Technocracy and Politics, Universidade Federal do Paraná, Curitiba, Brazil.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Garrido-Vergara, L. (2018). {\itshape Trajectories and Post-Career of Ministers in Chile}. Presentation delivered at the Research Seminar of the Institute of Public Affairs, Universidad de Chile, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Garrido-Vergara, L. (2018). {\itshape Trayectorias y poscarrera de ministros en Chile}. Presentation delivered at the Research Seminar of the Institute of Public Affairs, Universidad de Chile, Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Del Valle, N. (2017). {\itshape Media Pluralism and Digital Media Public Opinion}. Presentation delivered at the III Seminar of Pluralism and Mass Media in Chile, Universidad Alberto Hurtado, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Del Valle, N. (2017). {\itshape Pluralismo informativo y opinión pública en los medios digitales escritos}. Presentation delivered at the III Seminar of Pluralism and Mass Media in Chile: Challenges for the Democratisation of Communications, Universidad Alberto Hurtado, Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., Maillet, A., \& Olivares, A. (2017). {\itshape Revolving Door? Analysis of Public-Private Circulation in Chile}. Presentation delivered at the II Foreign Affairs Meeting, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B., Maillet, A., \& Olivares, A. (2017). {\itshape ?`Puerta giratoria? Análisis de la circulación público-privada en Chile}. Presentation delivered at the II Foreign Affairs Meeting, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Garrido-Vergara, L. (2017). {\itshape Socialisation Spaces and Ministerial Post-Career in the Chilean Transition and Post-Transition}. Presentation delivered at the IX Chilean Congress of Public Administration, Concepcion, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Garrido-Vergara, L. (2017). {\itshape Espacios de socialización, carrera y postcarrera ministerial en la transición y post-transición chilena}. Presentation delivered at the IX Chilean Congress of Public Administration, Concepcion, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., Maillet, A., \& Olivares, A. (2017). {\itshape Public-Private: An Especially Tenuous Frontier. Political-Professional Trajectories of the Chilean Authorities (2000-2014)}. Presentation delivered at the IX Latin American Congress of Political Science, Montevideo, Uruguay.}\vspace{1mm}

\item{González-Bustamante, B., Maillet, A., \& Olivares, A. (2017). {\itshape Público-privado: una frontera muy porosa. Trayectorias político-profesionales de autoridades en Chile (2000-2014)}. Presentation delivered at the IX Latin American Congress of Political Science (ALACIP), Montevideo, Uruguay.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Joignant, A. (2016). {\itshape Gaps and Elite-Mass Congruence in Chile and Uruguay}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Joignant, A. (2016). {\itshape Brechas y congruencia élite-ciudadanía en Chile y Uruguay}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

%% \item{Joignant, A., \& González-Bustamante, B. (2016). {\itshape Family Ties in the Senate and the Chamber of Deputies, Chile from 1989 to 2014}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

\item{Joignant, A., \& González-Bustamante, B. (2016). {\itshape Lazos familiares en el Senado y la Cámara de Diputados, Chile 1989-2014}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

%% \item{Barría, D., González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Analysis of the Academic Production on Open Government in Spanish}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

\item{Barría, D., González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Análisis de la producción académica en español sobre gobierno abierto}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

%% \item{Maillet, A., González-Bustamante, B., \& Olivares, A. (2016). {\itshape Revolving Door and Public-Private Circulation in Chile}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

\item{Maillet, A., González-Bustamante, B., \& Olivares, A. (2016). {\itshape Puerta giratoria y circulación público-privada en Chile}. Presentation delivered at the Seminar of Political, Economic, and Intellectual Elites: Gaps and Inequalities, Universidad de Santiago de Chile (USACH), Santiago, Chile.}\vspace{1mm}

%% \item{Barría, D., González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Open Government in Latin America and Spain: Analysis of the Scientific Production in Spanish}. Presentation delivered at the XXI CLAD International Congress on the State Reform and Public Administration, Santiago, Chile.}\vspace{1mm}

\item{Barría, D., González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Gobierno abierto en América Latina y España: Análisis de la producción en español sobre el tema}. Presentation delivered at the XXI CLAD International Congress on the State Reform and Public Administration, Santiago, Chile.}\vspace{1mm}

%% \item{Joignant, A., \& González-Bustamante, B. (2016). {\itshape Performance of Family Political Capital in the Chilean Post-Authoritarianism}. Presentation delivered at the XII Chilean Congress of Political Science, Pucon, Chile.}\vspace{1mm}

\item{Joignant, A., \& González-Bustamante, B. (2016). {\itshape Rendimiento del capital político familiar en el Chile postautoritario}. Presentation delivered at the XII Chilean Congress of Political Science, Pucon, Chile.}\vspace{1mm}

%% \item{Olivares, A., \& González-Bustamante, B. (2016). {\itshape Turnover and Ministerial Survival Before the Chilean Democratic Breakdown}. Presentation delivered at the XII Chilean Congress of Political Science, Pucon, Chile.}\vspace{1mm}

\item{Olivares, A., \& González-Bustamante, B. (2016). {\itshape Rotación y supervivencia ministerial antes del quiebre democrático en Chile}. Presentation delivered at the XII Chilean Congress of Political Science (ACCP), Pucon, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Barría, D. (2016). {\itshape Production Patterns of the Contemporary Knowledge in Social Sciences}. Presentation delivered at the IX Chilean Congress of Sociology, PreALAS Meeting, Talca, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Barría, D. (2016). {\itshape Lógicas de producción del conocimiento en las ciencias sociales contemporáneas}. Presentation delivered at the IX Chilean Congress of Sociology, PreALAS Meeting, Talca, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Co-Authorship Networks of the Political Science in Spanish (2010-2015)}. Presentation delivered at the IX Chilean Congress of Sociology, PreALAS Meeting, Talca, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Cisternas, C. (2016). {\itshape Redes de co-citación en la ciencia política en español (2010-2015)}. Presentation delivered at the IX Chilean Congress of Sociology, PreALAS Meeting, Talca, Chile.}\vspace{1mm}

%% \item{Cisternas, C., González-Bustamante, B., \& Barría, D. (2016). {\itshape Analysis of the Scientific Production in Spanish on Public Administration and Public Policies (2010-2012)}. Presentation delivered at the IX Chilean Congress of Sociology, PreALAS Meeting, Talca, Chile.}\vspace{1mm}

\item{Cisternas, C., González-Bustamante, B., \& Barría, D. (2016). {\itshape Análisis de la producción científica en español sobre administración pública y políticas públicas (2010-2012)}. Presentation delivered at the IX Chilean Congress of Sociology, PreALAS Meeting, Talca, Chile.}\vspace{1mm}

%% \item{Cisternas, C., \& González-Bustamante, B. (2016). {\itshape The CIEPLAN Monks: Intellectual Production and Co-Citation Networks During the Eighties}. Presentation delivered at the International Seminar of Social Sciences at the Crossroad: Latin American Intellectuals and Technocrats in Authoritarianism Times (1969-1990), Universidad Bernardo O'Higgins, Santiago, Chile.}\vspace{1mm}

\item{Cisternas, C., \& González-Bustamante, B. (2016). {\itshape Los monjes de CIEPLAN: producción intelectual y redes de citación durante los años ochenta}. Presentation delivered at the International Seminar on Social Sciences at the Crossroad: Latin American Intellectuals and Technocrats in Authoritarianism Times (1969-1990), Universidad Bernardo O'Higgins, Santiago, Chile.}\vspace{1mm}

\item{Alenda, S., \& González-Bustamante, B. (2016). {\itshape The Reshaping of the Political Right in Chile: Change and Continuity in the Perceptions of Social Inequalities}. Presentation delivered at the III ISA Forum of Sociology, Vienna, Austria.}\vspace{1mm}

\item{González-Bustamante, B. (2016). {\itshape Representation and Elite-Mass Congruence in Chile and Uruguay}. Presentation delivered at the XXXIV International Congress of the Latin American Studies Association (LASA), New York, United States.}\vspace{1mm}

\item{Alenda, S., González-Bustamante, B., \& Gartenlaub, A. (2016). {\itshape The Reshaping of the Political Right in Chile}. Presentation delivered at the XXXIV International Congress of the Latin American Studies Association (LASA), New York, United States.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Barría, D. (2015). {\itshape Stateness in the Chilean Post-Transition}. Presentation delivered at the VIII Chilean Congress of Public Administration, Rancagua, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Barría, D. (2015). {\itshape Demandas por más intervención estatal en el Chile postransicional}. Presentation delivered at the VIII Chilean Congress of Public Administration, Rancagua, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2015). {\itshape Media Consumption and Political Participation in Argentina, Chile, and Uruguay}. Presentation delivered at the XII Argentine Congress of Political Science, Mendoza, Argentina.}\vspace{1mm}

\item{González-Bustamante, B. (2015). {\itshape Consumo de medios y formas de participación política en Argentina, Chile y Uruguay}. Presentation delivered at the XII Argentine Congress of Political Science (SAAP), Mendoza, Argentina.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Olivares, A. (2015). {\itshape Trajectories of Ministers and Undersecretaries in Chile During Concertacion Governments (1990-2010)}. Presentation delivered at the VIII Latin American Congress of Political Science, Lima, Peru.}\vspace{1mm}

\item{González-Bustamante, B., \& Olivares, A. (2015). {\itshape Trayectorias de ministros y subsecretarios en Chile durante los gobiernos de la Concertación (1990-2010)}. Presentation delivered at the VIII Latin American Congress of Political Science (ALACIP), Lima, Peru.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Barría, D. (2015). {\itshape Stateness' Support in Chile from 2007 to 2014}. Presentation delivered at the VIII Latin American Congress of Political Science, Lima, Peru.}\vspace{1mm}

\item{González-Bustamante, B., \& Barría, D. (2015). {\itshape Estatalidad y tendencias estatistas en Chile (2007-2014)}. Presentation delivered at the VIII Latin American Congress of Political Science (ALACIP), Lima, Peru.}\vspace{1mm}

%% \item{Olivares, A., \& González-Bustamante, B. (2015). {\itshape Survival of Ministers and Undersecretaries in Chile (1990-2014)}. Presentation delivered at the International Congress of Elites and Leadership, Universidad  de Salamanca, Salamanca, Spain.}\vspace{1mm}

\item{Olivares, A., \& González-Bustamante, B. (2015). {\itshape Supervivencia de ministros y subsecretarios en Chile (1990-2014)}. Presentation delivered at the International Congress of Elites and Leadership in Times of Change, Universidad de Salamanca, Salamanca, Spain.}\vspace{1mm}

%% \item{Garrido-Vergara, L., \& González-Bustamante, B. (2015). {\itshape Elites and Political Reforms: The Chilean Case}. Presentation delivered at the International Congress of Elites and Leadership, Universidad  de Salamanca, Salamanca, Spain.}\vspace{1mm}

\item{Garrido-Vergara, L., \& González-Bustamante, B. (2015). {\itshape Las élites y las reformas políticas. El caso chileno}. Presentation delivered at the International Congress of Elites and Leadership in Times of Change, Universidad de Salamanca, Salamanca, Spain.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Ruiz, I. (2014). {\itshape Radiography to the Popular Election of Regional Councillors 2013}. Presentation delivered at the XI Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Ruiz, I. (2014). {\itshape Radiografía de la elección de consejeros regionales 2013}. Presentation delivered at the XI Chilean Congress of Political Science (ACCP), Santiago, Chile.}\vspace{1mm}

%% \item{Thezá, M., \& González-Bustamante, B. (2014). {\itshape Dimensions of the Young People Multi-Associative Participation in Chile}. Presentation delivered at the XI Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{Thezá, M., \& González-Bustamante, B. (2014). {\itshape Dimensiones de la participación multi-asociativa de los jóvenes en Chile}. Presentation delivered at the XI Chilean Congress of Political Science (ACCP), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2014). {\itshape Social Media and Malaise with Democracy in Argentina and Uruguay}. Presentation delivered at the VI Latin American Congress of Public Opinion, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2014). {\itshape Medios sociales y malestar con la democracia en Argentina y Uruguay}. Presentation delivered at the VI Latin American Congress of Public Opinion (WAPOR), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2014). {\itshape Media Consumption and Political Participation in the Southern Cone}. Presentation delivered at the VI Latin American Congress of Public Opinion, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2014). {\itshape Consumo de medios y participación política en el Cono Sur}. Presentation delivered at the VI Latin American Congress of Public Opinion (WAPOR), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2014). {\itshape Chilean Elections 2013: Incumbency, Campaign Spending, and Electoral Success in the Congress}. Presentation delivered at the Centre for Political Analysis and Research Colloquium, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2014). {\itshape Las elecciones chilenas 2013: Incumbencia, gasto y eficiencia electoral en la competencia al Congreso}. Presentation delivered at the Centre for Political Analysis and Research Colloquium (CAIP), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2013). {\itshape Digital Independent Media in Chile}. Presentation delivered at the Seminar of Mass Media Pluralism in Chile, Centre for Political Analysis and Research, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2013). {\itshape Medios independientes y digitales en Chile}. Presentation delivered at the Seminar of Mass Media Pluralism in Chile, Centre for Political Analysis and Research (CAIP), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2013). {\itshape The Governmental Political Elite in Chile}. Presentation delivered at the Seminar of Elites Studies in Chile, Centre for Political Analysis and Research, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2013). {\itshape La élite política gubernamental (1990-2010)}. Presentation delivered at the Seminar of Elites Studies in Chile, Centre for Political Analysis and Research (CAIP), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2013). {\itshape Regional Barometer 2013: Regional Public Opinion in Bio-Bio, Los Lagos, and Los Rios}. Presentation delivered at the VI Chilean Congress of Public Administration, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2013). {\itshape Barómetro Regional 2013, opinión pública regional en Biobío, Los Lagos y Los Ríos}. Presentation delivered at the VI Chilean Congress of Public Administration, Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2013). {\itshape Online Social Networks and Political Behaviour in Chile: Twitter as a Measure of Public Opinion During the Primary Presidential Elections in 2013}. Presentation delivered at the II International Congress of Political Communication, Toluca, Mexico.}\vspace{1mm}

\item{González-Bustamante, B. (2013). {\itshape Redes sociales y comportamiento político en Chile: Twitter como indicador de opinión pública en las primarias presidenciales de 2013}. Presentation delivered at the II International Congress of Political Communication (ALICE), Toluca, Mexico.}\vspace{1mm}

%% \item{Chávez, D., \& González-Bustamante, B. (2013). {\itshape Partisanship Identification and Disaffection in Ecuador}. Presentation delivered at the XXIX Latin American Congress of Sociology, Santiago, Chile.}\vspace{1mm}

\item{Chávez, D., \& González-Bustamante, B. (2013). {\itshape Identificación y desafección partidaria en el Ecuador de hoy}. Presentation delivered at the XXIX Latin American Congress of Sociology (ALAS), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Soto, D. (2013). {\itshape Digital Independent Media in Chile: Democratisation in the Media System?} Presentation delivered at the VII Latin American Congress of Political Science, Bogota, Colombia.}\vspace{1mm}

\item{González-Bustamante, B., \& Soto, D. (2013). {\itshape Medios independientes y digitales en Chile: ?`Democratización para el sistema de medios de comunicación?} Presentation delivered at the VII Latin American Congress of Political Science (ALACIP), Bogota, Colombia.}\vspace{1mm}

%% \item{González-Bustamante, B. (2013). {\itshape Trajectories and Patterns of the Ministers' Political Careers in Chile}. Presentation delivered at the XI Argentine Congress of Political Science, Parana, Argentina.}\vspace{1mm}

\item{González-Bustamante, B. (2013). {\itshape Trayectorias y patrones de carreras políticas de ministros en Chile (1990-2010)}. Presentation delivered at the XI Argentine Congress of Political Science (SAAP), Parana, Argentina.}\vspace{1mm}

%% \item{González-Bustamante, B. (2013). {\itshape Twitter as Public Opinion Measure: The Bachelet's Arrival to the Chilean Presidential Campaign 2013}. Presentation delivered at the XI Argentine Congress of Political Science, Parana, Argentina.}\vspace{1mm}

\item{González-Bustamante, B. (2013). {\itshape Twitter como indicador de opinión pública: El arribo de Bachelet a la presidencial chilena 2013}. Presentation delivered at the XI Argentine Congress of Political Science (SAAP), Parana, Argentina.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Valenzuela, P. (2012). {\itshape Representation Gaps and Incorporated Capital: The Case of the Concertacion's Political Youths}. Presentation delivered at the VII Chilean Congress of Sociology, PreALAS Meeting, Pucon, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Valenzuela, P. (2012). {\itshape Brechas de representación y capital incorporado: El caso de las juventudes concertacionistas}. Presentation delivered at the VII Chilean Congress of Sociology, PreALAS Meeting, Pucon, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B. (2012) {\itshape The Transversal Political Party Thesis: Beyond the Myth}. Presentation delivered at the X Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2012) {\itshape La tesis del partido transversal: Más allá del mito}. Presentation delivered at the X Chilean Congress of Political Science (ACCP), Santiago, Chile.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Chávez, D. (2012). {\itshape Bolivian Inter-Cultural Democracy: Advances and Challenges}. Presentation delivered at the X Chilean Congress of Political Science, Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B., \& Chávez, D. (2012). {\itshape Democracia intercultural boliviana: Avances y desafíos}. Presentation delivered at the X Chilean Congress of Political Science (ACCP), Santiago, Chile.}\vspace{1mm}

\item{González-Bustamante, B. (2012). {\itshape Lights and Shadows of the Public Management Professionalisation in the Chilean Case}. Presentation delivered at the XXII World Congress of Political Science (IPSA), Madrid, Spain.}\vspace{1mm}

\item{Ríos, T. E., \& González-Bustamante, B. (2012). {\itshape Media System and Democratisation in the Chilean Case}. Presentation delivered at the XXII World Congress of Political Science (IPSA), Madrid, Spain.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Henríquez, G. (2012). {\itshape The Role of the Online Social Networks in the Last Chilean Presidential Campaign}. Presentation delivered at the I International Congress of Political Communication, Madrid, Spain.}\vspace{1mm}

\item{González-Bustamante, B., \& Henríquez, G. (2012). {\itshape El rol de las redes sociales en la última campaña presidencial chilena}. Presentation delivered at the I International Congress of Political Communication (ALICE), Madrid, Spain.}\vspace{1mm}

%% \item{González-Bustamante, B. (2012). {\itshape Research Notes and Reflections on the Study of Elites in Chile}. Presentation delivered at the VI Latin American Congress of Political Science, Quito, Ecuador.}

\item{González-Bustamante, B. (2012). {\itshape Notas y reflexiones sobre el estudio de las élites en Chile}. Presentation delivered at the VI Latin American Congress of Political Science (ALACIP), Quito, Ecuador.}

\end{benumerate}

\end{publications}