### 2024-06-24 sections/teaching.tex
Update teaching.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% Leiden University
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Relevant Teaching Experience}

\begin{publications}

\begin{itemize}
\item{\small School of Business, {\bfseries Universidad Mayor}, Santiago, CHL}
\begin{itemize}
%% \item[$\circ$]{\small Machine Learning in Public Policy and Economics (workshop, Prof. Carla Cisternas), spring 2021}
\item[$\circ$]{\small Machine Learning in Public Policy and Economics (workshop), spring 2021}
\end{itemize}
\item{\small Department of Politics and International Relations, {\bfseries University of Oxford}, Oxford, GBR}
\begin{itemize}
%% \item[$\circ$]{\small Comparative Government (undergraduate tutorial, Prof. Petra Schleiter), Hilary 2020}
\item[$\circ$]{\small Comparative Government (undergraduate tutorial), Hilary 2020}
\end{itemize}
\item{\small Faculty of Adminstration and Economics, {\bfseries Universidad de Santiago de Chile}, Santiago, CHL}
\begin{itemize}
\item[$\circ$]{\small Political Science I (undergraduate lecture), spring 2014, 2018, fall 2015-2019}
\item[$\circ$]{\small Political Science II (undergraduate lecture), spring 2015-2018}
\item[$\circ$]{\small Econometrics for Public Policies (undergraduate lecture), fall 2018}
\item[$\circ$]{\small Workshop for Research Design (graduate seminar), spring 2017-2018}
\item[$\circ$]{\small Research Methodology (undergraduate lecture), spring 2014-2015, fall 2015-2017}
\item[$\circ$]{\small Introduction to Quantitative Methods (undergraduate lecture), fall 2017}
\end{itemize}
\item{\small School of Government and Public Management, {\bfseries Universidad de Chile}, Santiago, CHL}
\begin{itemize}
\item[$\circ$]{\small Study of Political Phenomena (undergraduate lecture), fall 2016}
\item[$\circ$]{\small Political Analysis (undergraduate lecture), fall 2014}
%% \item[$\circ$]{\small Elite Theory and Democracy (undergraduate lecture, Prof. Alejandro Olivares), spring 2013-2014}
\item[$\circ$]{\small Elite Theory and Democracy (undergraduate lecture), spring 2013-2014}
\end{itemize}
\end{itemize}

\vspace{1mm}
\end{publications}
\pagebreak