### 2024-06-24 short/sections/future-papers.tex
Create future-papers.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% Leiden University
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Working Papers}

\begin{publications}

\begin{itemize}

\item{Kantorowicz, J., \& González-Bustamamante, B. (2024, {\itshape forthcoming}). {\itshape The Changing Concepts of the Rule of Law}. Presentation delivered at the 6th Annual COMPTEXT Conference, Amsterdam, Netherlands.}\vspace{1mm}

\item{González-Bustamamante, B., \& Kantorowicz, J. (2024, {\itshape forthcoming}). {\itshape Multilingual Word Embeddings for Unravelling the Rule of Law in European Countries}. Presentation delivered at the 6th Annual COMPTEXT Conference, Amsterdam, Netherlands.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Machine Learning and Political Events: Application of a Semi-Supervised Approach to Produce a Dataset on Presidential Cabinets}. Doctoral thesis chapter, University of Oxford.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Resignation Calls, Reallocations and Individual Ministerial Terminations in Presidential Democracies}. Doctoral thesis chapter, University of Oxford.}\vspace{1mm}

\item{González-Bustamante, B. (2023). {\itshape Corrective Effect of Ministerial Terminations on Presidential Approval}. Doctoral thesis chapter, University of Oxford.}\vspace{1mm}

\end{itemize}

\end{publications}

%% \pagebreak