### 2024-07-29 short/sections/summary.tex
Update summary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% Leiden University
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{3mm}
\cvsection{Summary}

\begin{cvparagraph}

\textcolor{black}{I am a post-doctoral researcher in Computational Social Science at the Institute of Public Administration at the Faculty of Governance and Global Affairs at Leiden University, Netherlands. I hold a DPhil (PhD) in Politics from the University of Oxford, United Kingdom. Before my DPhil, I earned an MA (1st) in Political Science and a BA (2:1) in Government, both from the Universidad de Chile.}

\textcolor{black}{My research interests lie in the intersection of comparative politics and government, focusing mainly on cabinets, political regimes and civil services. Methodologically, my interests rely on the application of quantitative text analysis, machine learning methods and causal inference strategies in the comparative politics field. In the last years, my work has been published in {\itshape The International Journal of Press/Politics}, {\itshape World Development}, {\itshape Government and Opposition}, {\itshape The British Journal of Politics and International Relations}, {\itshape Bulletin of Latin American Research} and elsewhere.}

\textcolor{black}{During my post-doctorate at Leiden, I am working on two research projects funded by the Dutch Research Council (NWO) related to how the rule of law has changed over time and automated text analysis of investment data to study the effects of governance on sustainable investments. Specifically, I employ Natural Language Processing (NLP) and text-as-data techniques to find patterns and topics within big text data.}
\vspace{1mm}
\end{cvparagraph}