### 2024-09-24 sections/awards.tex
Update awards.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Relevant Honours \& Awards}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Early Career Scholar Award}
{Best early career work and travel support for the BigSurv23 Conference, about {\pounds} 525.}
{Quito, ECU}
{2023}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{ECPR Joint Sessions of Workshops Grant}
{Travel support for the European Consortium for Political Research (ECPR) Joint Sessions of Workshops, about {\pounds} 1 350.}
{Toulouse, FRA}
{2023}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{SLAS Postgraduate and Postdoctoral Research Award and Conference Grant} 
{Research support grant for my DPhil (PhD) project and travel support for the Society for Latin American Studies (SLAS) Annual Conference, about {\pounds} 1 400.}
{Bath, GBR}
{2022}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{St Hilda's College Muriel Wise Fund and Travel and Book Grants} 
{Research support grant for my DPhil (PhD) project and allowances for participation in conferences and purchase of books, about {\pounds} 2 255.} 
{Oxford, GBR}
{2022}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Becas Chile Doctoral Scholarship Abroad} 
{Tuition fees, transport and living expenses for pursuing a DPhil (PhD) at the University of Oxford (ANID/PFCHA/72200340), about {\pounds} 160 000.}
{Oxford, GBR}
{2019}
\end{cvhonors}

\vspace{1mm}

%% \begin{cvhonors}
%% \cvhonor
%% {LASA Travel Grant} 
%% {Travel support for the XXXIV International Congress of the Latin American Studies Association (LASA/40071), {\pounds} 400.}
%% {New York, USA}
%% {2016}
%% \end{cvhonors}

%% \vspace{1mm}

%% \begin{cvhonors}
%% \cvhonor
%% {CESS Summer School Scholarship}
%% {Financial support for the Experimental Summer School Methods at the Nuffield-USACH Centre for Experimental Social Sciences (CESS), {\pounds} 200}
%% {Santiago, CHL}
%% {2016}
%% \end{cvhonors}

%% \vspace{1mm}

\begin{cvhonors}
\cvhonor
{ALACIP Methodological School Prize}
{Best methodological work at the Latin American Political Association (ALACIP) School.}
{Lima, PER}
{2015}
\end{cvhonors}

\vspace{1mm}

%% \begin{cvhonors}
%% \cvhonor
%% {ACCP Winter School Scholarship}
%% {Financial support for the Winter School in Methods and Data Analysis at the Universidad Cat\'olica de Uruguay, awarded by the Chilean Political Science Association (ACCP), {\pounds} 300}
%% {Montevideo, URY}
%% { 2015}
%% \end{cvhonors}

%% \vspace{1mm}

%% \begin{cvhonors}
%% \cvhonor
%% {PUC Summer School Scholarship}
%% {Financial support for the Summer School in Mixed Methods at the Pontificia Universidad Cat\'olica de Chile (PUC), {\pounds} 170}
%% {Santiago, CHL}
%% {2014} 
%% \end{cvhonors}

%% \vspace{1mm}

\begin{cvhonors}
\cvhonor
{Comunicracia Prize}
{Best paper of the political communication area at the XI Argentine Congress of Political Science (double-blind peer-review).} %% {\pounds} 170
{Parana, ARG}
{2013}
\end{cvhonors}

\vspace{1mm}

%% \begin{cvhonors}
%% \cvhonor
%% {Highest Rating MA Dissertation}
%% {The highest rating and the congratulations from the jury at the Universidad de Chile}
%% {Santiago, Chile}
%% {2013}
%% \end{cvhonors}

%% \vspace{1mm}

\begin{cvhonors}
\cvhonor
{FONDECYT Thesis Writing Grant}
{Financial support for MA fieldwork (Prof. Gonzalo Delamaza, Supervisor, ANID/FONDECYT/1100877), about {\pounds} 1 150.}
{Santiago, CHL}
{2012}
\end{cvhonors}

\vspace{1mm}