### 2024-09-24 sections/projects.tex
Create projects.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Relevant Research Projects}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Making Finance Sustainable? A Comparative Perspective on the Politics of Investment}
{NWO VIDI SWG Project VI.Vidi.211.226, Research Associate (Natascha van der Zwan, Principal Investigator), about {\pounds} 800 000. Period: 2024-2026.}
{Dutch Research Agency}
{2024}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Validación de un estándar de referencia para medir toxicidad e incivilidad política en la esfera digital}
{Internal Project 2024 at the Faculty of Administration and Economics at the Universidad Diego Portales (UDP), Principal Investigator, about {\pounds} 1 500. Period: 2024-2025.}
{UDP}
{2024}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Large Language Models (LLMs) to Identify Toxicity in the Digital Sphere during Protest Events in Latin America}
{OpenAI Academic Programme, Principal Investigator, about {\pounds} 5 000. Period: 2024.}
{OpenAI}
{2024}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Unravelling the Rule of Law: A Machine Learning Approach}
{NWO XS Project 406.XS.01.013, Research Associate (Jaroslaw Kantorowicz, Principal Investigator), about {\pounds} 45 000. Period: 2023-2024.}
{Dutch Research Agency}
{2023}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Patrones territoriales de gobierno electrónico y abierto en los municipios chilenos}
{Internal Project USA 2156 POC2022\_FAE2 at the Universidad de Santiago de Chile (USACH), Principal Investigator, about {\pounds} 1 700. Period: 2022.}
{USACH}
{2022}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Sistema político y visibilidad de actores en el debate público articulado por la prensa de élite: 1990-2018}
{XVI Internal Competition for the Promotion of Research 2019 at the Universidad Alberto Hurtado (UAH), Resarch Associate (Ximena Orchard, Principal Investigator), about {\pounds} 3 000. Period: 2019.}
{UAH}
{2019}
\end{cvhonors}

\vspace{1mm}