### 2024-09-24 sections/reviewer.tex
Update reviewer.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Peer Reviewer, Expert \& Committees}

\begin{publications}

\begin{itemize}
\item{\small Varieties of Democracy Project, University of Gothenburg, SWE, 2014, 2018-2020, 2024-2025}
\item{\small Moderator in {\itshape SocArXiv}, University of Maryland, USA, {\itshape since} 2023}
\item{\small Experts Network for Transparency, Pan American Development Foundation, ECU, {\itshape since} 2022}
\item{\small Editor in {\itshape Tufte Working Papers} (ISSN 2735-6043), Training Data Lab, CHL, {\itshape since} 2020}
\item{\small Associate Editor in {\itshape Revista de Gestión Pública} (ISSN 0719-1820), Universidad de Valparaíso, CHL, {\itshape since} 2016}
\item{\small Reviewer in {\scshape WoS-SSCI}, {\scshape Scopus}, \& {\scshape WoS-ESCI} journals, {\itshape since} 2013}
\item{\small Reviewer in {\scshape SciELO} \& {\scshape Latindex} journals, {\itshape since} 2013}
\item{\small Times Higher Education World University Rankings, {\scshape Elsevier}, NLD, 2018-2019, 2022-2024}
\item{\small Master in Economic History, Universidad de Santiago de Chile (USACH), CHL, 2023-2024}
\item{\small Master in Management and Public Policy, Universidad de Santiago de Chile (USACH), CHL, 2022-2024}
\item{\small Global Academic Reputation Survey, Clarivate Analytics, USA, 2020}
\item{\small Sectoral Commission for Research, Universidad de la República, URY, 2016, 2018}
\end{itemize}

\vspace{1mm}
%% \pagebreak
\end{publications}