### 2024-09-24 sections/summary.tex
Update summary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{3mm}
\cvsection{Summary}

\begin{cvparagraph}

\textcolor{black}{I am a post-doctoral researcher in Computational Social Science and a lecturer in Governance and Development at the Institute of Public Administration at the Faculty of Governance and Global Affairs at Leiden University, Netherlands. I hold a DPhil (PhD) in Politics from the University of Oxford, United Kingdom. Before my DPhil, I earned an MA (1st) in Political Science and a BA (2:1) in Public Administration, both from the Universidad de Chile.}

\textcolor{black}{In addition, in Chile, I am a faculty member at the Faculty of Administration and Economics at the Universidad Diego Portales and a research associate at the Training Data Lab, a research group focused on data mining, econometric modelling and AI machine learning in social sciences.}

\textcolor{black}{My research interests lie in the intersection of comparative politics and government, focusing mainly on cabinets, political regimes and civil services. Methodologically, my interests rely on the application of quantitative text analysis, machine learning methods and causal inference strategies in the comparative politics field. In the last years, my work has been published in {\itshape The International Journal of Press/Politics}, {\itshape World Development}, {\itshape Government and Opposition}, {\itshape The British Journal of Politics and International Relations}, {\itshape Bulletin of Latin American Research} and elsewhere.}
\vspace{1mm}
\end{cvparagraph}