### 2025-01-15 short/sections/summary.tex
Update summary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% Leiden University
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{3mm}
\cvsection{Summary}

\begin{cvparagraph}

\textcolor{black}{I am a post-doctoral researcher in Computational Social Science at the Institute of Public Administration, Faculty of Governance and Global Affairs, Leiden University, the Netherlands. I hold a DPhil (PhD) in Politics from the University of Oxford, UK, and prior to that, I completed an MA (1st) in Political Science and a BA (2:1) in Government at the Universidad de Chile.}

\textcolor{black}{My research, which initially focused on comparative politics and government, particularly cabinets, political regimes and civil services, has now expanded to explore the emerging field of digital governance and other policy domains. My work has appeared in {\itshape Social Science Computer Review}, {\itshape World Development}, {\itshape Government and Opposition}, {\itshape The International Journal of Press/Politics}, {\itshape The British Journal of Politics} and elsewhere.}

\textcolor{black}{During my current postdoctoral appointment, I am mainly contributing to an NWO-funded project on automated text analysis of investment data, investigating how governance structures shape sustainable investments. Specifically, I use Natural Language Processing (NLP) and text-as-data methods to identify patterns and topics within large-scale textual corpora. In addition, I teach the Institutions of Governance and Development course at Leiden University College.}
\vspace{1mm}
\end{cvparagraph}