### 2025-02-08 short/sections/training.tex
Create training.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Relevant Certification \& Training}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{KODAQS Data Quality Academy Certificate Program, University of Mannheim} 
{Data quality dimensions, indicators and metrics to assess data quality, tools and data quality, data management procedures, reproducibility} 
{Mannheim, DEU}
{2025}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{ODISSEI Summer Institute in Computational Social Science} 
{Network analysis, machine learning, benchmarking and ethics} 
{Rotterdam, NLD}
{2024}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Introduction to Learning and Teaching at Oxford} 
{Foundations for inclusive teaching and strategies to support efficient and effective preparation for teaching} 
{Oxford, GBR}
{2023}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Oxford|Berlin Summer School on Open Research} 
{Training in biases in research, reproducibility, publishing, open data, open materials, meta-research and ethical conduct of research} 
{Oxford, GBR}
{2021}
\end{cvhonors}

\vspace{1mm}
%% \pagebreak