### 2025-02-08 sections/references.tex
Update references.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Referees}

\begin{publications}

\begin{itemize}
\item[]{\small {\bfseries Natascha van der Zwan}, Associate Professor, Leiden University, NLD}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:n.a.j.van.der.zwan@fgga.leidenuniv.nl}{\textcolor{blue}{n.a.j.van.der.zwan@fgga.leidenuniv.nl}}}} \vspace{1.5mm}
\item[]{\small {\bfseries Jaroslaw Kantorowicz}, Associate Professor, Leiden University, NLD}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:j.j.kantorowicz@fgga.leidenuniv.nl}{\textcolor{blue}{j.j.kantorowicz@fgga.leidenuniv.nl}}}} \vspace{1.5mm}
\item[]{\small {\bfseries Petra Schleiter}, Professor of Comparative Politics, University of Oxford, GBR}
\item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:petra.schleiter@politics.ox.ac.uk}{\textcolor{blue}{petra.schleiter@politics.ox.ac.uk}}}} \vspace{1.5mm}
%% \item[]{\small {\bfseries David Doyle}, Head of Department and Professor of Politics, University of Oxford, GBR}
%% \item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:david.doyle@politics.ox.ac.uk}{\textcolor{blue}{david.doyle@politics.ox.ac.uk}}}} \vspace{1.5mm}
%% \item[]{\small {\bfseries Ryan E. Carlin}, Professor of Political Science, Georgia State University, USA}
%% \item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small %% \href{mailto:rcarlin@gsu.edu}{\textcolor{blue}{rcarlin@gsu.edu}}}} \vspace{1.5mm}
%% \item[]{\small {\bfseries Jaroslaw Kantorowicz}, Associate Professor, Leiden University, NLD}
%% \item[]{\vspace{-1mm}\textcolor{oxfordblue}{\faEnvelope} \hspace{0.1mm} {\small \href{mailto:j.j.kantorowicz@fgga.leidenuniv.nl}{\textcolor{blue}{j.j.kantorowicz@fgga.leidenuniv.nl}}}} \vspace{1.5mm}
\end{itemize}

\end{publications}