### 2025-02-12 short/sections/summary.tex
Update summary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% Leiden University
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{3mm}
\cvsection{Summary}

\begin{cvparagraph}

\textcolor{black}{I am a post-doctoral researcher in Computational Social Science at the Institute of Public Administration, Faculty of Governance and Global Affairs, Leiden University, the Netherlands. I hold a DPhil (PhD) in Politics from the University of Oxford, UK, and prior to that, I completed an MA (1st) in Political Science and a BA (2:1) in Government at the Universidad de Chile.}

\textcolor{black}{From an earlier focus on cabinets and civil service, I have developed an active research agenda on how emerging technologies ---including AI--- shape policymaking and the organisation of bureaucracies. My work has appeared in {\itshape Social Science Computer Review}, {\itshape World Development}, {\itshape Government and Opposition}, {\itshape The International Journal of Press/Politics}, {\itshape The British Journal of Politics and International Relations} and elsewhere.}

\textcolor{black}{During my current postdoctoral appointment, I am mainly contributing to an NWO-funded project on automated text analysis of investment data, investigating how governance structures shape sustainable investments. Specifically, I use Natural Language Processing (NLP) and text-as-data methods to identify patterns and topics within large-scale textual corpora. In addition, I teach the Institutions of Governance and Development course at Leiden University College.}
\vspace{1mm}
\end{cvparagraph}