### 2025-06-06 sections/languages.tex
Update languages.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Languages \& Software}

\begin{cvskills}

\cvskill
{Languages} 
{\textcolor{black}{Spanish (native speaker), English (fluent), Portuguese (advanced reading skills), Dutch,}}

\cvskill
{} 
{\textcolor{black}{French and Italian (elemental reading skills)}} \\

\cvskill
{Programming \&} 
{\textcolor{black}{Python, R, Git, Stata, {\scshape spss/pspp}, Gephi, {\scshape ucinet}, NetDraw, Tosmana, ArcGIS, {\scshape atlas}.ti}, {\scshape HPoP},}

\cvskill
{Software} 
{\textcolor{black}{Zotero, JabRef, {\LaTeX}, Bib{\TeX}, MS Office, WordPress, Moodle, {\scshape java}, {\scshape html}, {\scshape css}, {\scshape xhtml}}} \\

\end{cvskills}