### 2025-06-06 short/sections/future-papers.tex
Update future-papers.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% Leiden University
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Papers Under Review}

\hspace{2.25mm}{\footnotesize $^\star$ {\itshape These manuscripts are in the first round of peer review or R\&R}}

\begin{publications}

\begin{benumerate}{6}

\item{Orchard, X., González-Bustamante, B., \& Cisternas, C. (2025, {\itshape second round}). Amplification, Contestation, and Toxic Interactions with News: Testing the Effect of a Protest Cycle in Digital News Participation. Preprint upon request.}\vspace{1mm}

\item{DiGiuseppe, M., Aspide, A., Kantorowicz, J., \& González-Bustamante, B. (2025, {\itshape second round}). Putting the Public Back in Public Debt: Citizen Narratives on Public Debt Burdens. Preprint upon request.}\vspace{1mm}

\item{Kantorowicz, J., \& González-Bustamante, B. (2024, {\itshape first round}). Exploring the Rule of Law in Academic Discourse: A Comprehensive Bibliometric Analysis. OSF preprint. DOI: \href{https://doi.org/10.31219/osf.io/sgf8y}{\textcolor{blue}{10.31219/osf.io/sgf8y}}.}\vspace{1mm}

\item{González-Bustamante, B. (2024, {\itshape first round}). Benchmarking LLMs in Political Content Text-Annotation: Proof-of-Concept with Toxicity and Incivility Data. arXiv preprint. DOI: \href{https://doi.org/10.48550/arXiv.2409.09741}{\textcolor{blue}{10.48550/arXiv.2409.09741}}.}\vspace{1mm}

\item{González-Bustamante, B. (2024, {\itshape first round}). Resignation Calls, Reallocations and Individual Ministerial Terminations in Presidential Democracies. Preprint upon request.}\vspace{1mm} %% \faBriefcase\ {\bfseries Job market paper}.}\vspace{1mm}

\item{González-Bustamante, B., \& Kantorowicz, J. (2024, {\itshape first round}). Cross-Lingual Word Embeddings for Unravelling the Rule of Law in European Countries. OSF preprint. DOI: \href{https://doi.org/10.31219/osf.io/m65re}{\textcolor{blue}{10.31219/osf.io/m65re}}.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Rivera, S. (2024, {\itshape first round}). Toxicity and Digital Incivility during Political Deliberation: The Case of the Constitutional Convention in Chile. Preprint upon request.}\vspace{1mm}

\end{benumerate}

\end{publications}

%% \pagebreak