### 2025-09-08 spanish/sections/working-papers.tex
Update working-papers.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in Spanish
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% \cvsection{Proyectos de recopilación de datos}

%% \begin{publications}

%% \begin{benumerate}{5}

%% \item{González-Bustamante, B., \& Aguilar, D. (2023). Data Set on Local Government Indicators in Chile (Version 0.20.15 -- Late Sky). Dataset, versión previa al lanzamiento, University of Oxford, Universidad de Santiago de Chile (USACH) y Training Data Lab. {\scshape doi:} \href{https://doi.org/10.5281/zenodo.7568387}{\textcolor{blue}{10.5281/zenodo.7568387}}.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Olivares, A. (2023). Data Set on Chilean Ministers (1990-2014) (Version 3.3.7 -- Soft Dew). Dataset, University of Oxford, Universidad de Santiago de Chile (USACH), Universidad Mayor y Training Data Lab. {\scshape doi:} \href{https://doi.org/10.5281/zenodo.5744536}{\textcolor{blue}{10.5281/zenodo.5744536}}.}\vspace{1mm}

%% \item{González-Bustamante, B., \& Olivares, A. (2023). Data Set on Chilean Undersecretaries (1990-2022) (Version 2.1.4 -- Wandering Grass). Dataset, University of Oxford, Universidad de Santiago de Chile (USACH), Universidad Mayor y Training Data Lab. {\scshape doi:} \href{https://doi.org/10.5281/zenodo.5715384}{\textcolor{blue}{10.5281/zenodo.5715384}}.}\vspace{1mm}

%% \item{González-Bustamante, B., Olivares, A., Cisternas, C., \& Cuevas, R. (2023). Chilean Political Science Ranking. Dataset, University of Oxford, Universidad de Santiago de Chile (USACH) y Training Data Lab. {\scshape doi:} \\ \href{http://doi.org/10.17605/OSF.IO/C8PRA}{\textcolor{blue}{10.17605/OSF.IO/C8PRA}}.}\vspace{1mm}

%% \item{González-Bustamante, B., Astete, M., \& Orvenes, B. (2021). Survival of the Senior Civil Servants in the Chilean Executive Branch (Version 1.13.19 -- Bold Lab). Dataset, versión previa al lanzamiento bajo acceso restringido, University of Oxford, Universidad de Santiago de Chile (USACH) and Training Data Lab. {\scshape doi:} \href{https://doi.org/10.5281/zenodo.8115596}{\textcolor{blue}{10.5281/zenodo.8115596}}.}\vspace{1mm}

%% \end{benumerate}

%% \end{publications}
%% \pagebreak

\cvsection{Documentos de trabajo y reseñas}

\begin{publications}

\begin{benumerate}{8}

\item{González-Bustamante, B. (2025). Charting Reproducibility and Performance: LLMs in Multilingual Toxic Speech Detection. Paper presented at the LLM Pre-Conference Workshop, European Political Science Association (EPSA), Universidad Carlos III de Madrid, Madrid, Spain, June 25. URL: \href{https://talkingtomachines.org/large-language-models-workshop/}{\textcolor{blue}{https://talkingtomachines.org}}.}\vspace{1mm}

\item{González-Bustamante,B. (2024). Benchmarking LLMs in Political Content Text-Annotation: Proof-of-Concept with Toxicity and Incivility Data. Paper presented at the 8th Monash-Warwick-Zurich Text-as-Data Workshop, September 16-17. arXiv: \href{https://doi.org/10.48550/arXiv.2409.09741}{\textcolor{blue}{10.48550/arXiv.2409.09741}}.}\vspace{1mm}

\item{González-Bustamante, B., \& Kantorowicz, J. (2024). Cross-Lingual Word Embeddings for Unravelling the Rule of Law in European Countries. Preprint. OSF: \href{https://doi.org/10.31219/osf.io/m65re}{\textcolor{blue}{10.31219/osf.io/m65re}}.}\vspace{1mm}

\item{Kantorowicz, J., \& González-Bustamante, B. (2024). Exploring the Rule of Law in Academic Discourse: A Comprehensive Bibliometric Analysis. Preprint. OSF: \href{https://doi.org/10.31219/osf.io/sgf8y}{\textcolor{blue}{10.31219/osf.io/sgf8y}}.}\vspace{1mm}

\item{Maillet, A., González-Bustamante, B., \& Olivares, A. (2016). ¿Puerta giratoria? Análisis de la circulación público-privada en Chile (2000-2014). {\itshape Serie Documentos de Trabajo PNUD-Desigualdad}, 7, 1--40. DOI: \\ \href{http://doi.org/10.13140/RG.2.2.25510.42566}{\textcolor{blue}{10.13140/RG.2.2.25510.42566}}.}\vspace{1mm}

\item{González-Bustamante, B., \& Olivares, A. (2016). A rotatividade de subsecretários dos ministérios no Chile (1990-2014). {\itshape Newsletter. Observat\'orio de elites pol\'iticas e sociais do Brasil. NUSP/UFPR, 3}(5), 1--19. URL: \\ \href{https://www.researchgate.net/publication/321993740_A_rotatividade_de_subsecretarios_dos_ministerios_no_Chile_1990-2014}{\textcolor{blue}{www.researchgate.net}}.}\vspace{1mm}

\item{González-Bustamante, B., \& Soto, D. (2016). Pluralismo en los medios digitales escritos en Chile, una exploración de casos clave. {\itshape Serie Política, Gestión y Políticas Públicas}, 2, 1--26. DOI: \href{http://doi.org/10.13140/RG.2.2.33899.03362/1}{\textcolor{blue}{10.13140/RG.2.2.33899.03362/1}}.}\vspace{1mm}

\item{González-Bustamante, B. (2015). Programas de Naciones Unidas para el Desarrollo (PNUD). (2014). Ciudadanía Política: Voz y Participación Ciudadana en América Latina. Buenos Aires: Siglo Veintiuno Editores. {\itshape Revista de Gestión Pública, IV}(1), 143--149. DOI: \href{https://doi.org/10.22370/rgp.2015.4.1.2240}{\textcolor{blue}{10.22370/rgp.2015.4.1.2240}}.}

\end{benumerate}

\end{publications}
\pagebreak