### 2025-10-28 short/sections/summary_associate.tex
Create summary_associate.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% Leiden University
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{3mm}
\cvsection{Summary}

\begin{cvparagraph}

\textcolor{black}{I hold dual appointments as an Associate Professor of Public Administration at Universidad Diego Portales and as a Postdoctoral Researcher in Computational Social Science at Leiden University’s Institute of Public Administration, Faculty of Governance and Global Affairs. I hold a DPhil (PhD) in Politics from the University of Oxford and previously earned an MA in Political Science and a BA in Public Administration and Government from the Universidad de Chile.}

\textcolor{black}{My work bridges the fields of comparative politics, government and computational social science. I build large-scale text-as-data pipelines, deploy AI and machine learning models, and apply causal inference strategies. My recent peer-reviewed articles have appeared in {\itshape Social Science Computer Review}, {\itshape Public Opinion Quarterly}, {\itshape World Development}, {\itshape Artificial Intelligence and Law}, {\itshape Government and Opposition}, among others. I regularly present at IPSA, ECPR, EPSA and COMPTEXT.}

\textcolor{black}{I am currently working on an NWO-funded project that uses Large Language Models (LLMs) to analyse global investment prospectuses and reveal how governance arrangements shape sustainable finance flows. In parallel, I contribute to COST Action CA22150 on executive-bureaucratic careers, lead the Enlace-Inserción UDP 2025-2026 project “Unpacking the Unpredictable: Using NLP and LLMs to Examine Cabinet Politics and Responses to Stochastic Events in Presidential Democracies,” and am a Research Leader at the Oxford Computational Political Science Group.}
\vspace{1mm}
\end{cvparagraph}