### 2025-10-28 short/sections/projects_associate.tex
Create projects_associate.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% CV-XeLaTeX in English
%% Bastián González-Bustamante
%% University of Oxford
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX

%% Based on the following repositories:
%% Awesome CV LaTeX Template for CV/Resume
%% https://github.com/posquit0/Awesome-CV
%% Bastián Gozález-Bustamante's CV-LaTeX Template
%% https://github.com/bgonzalezbustamante/CV-LaTeX
%% Carla Cisternas' CV-LuaLaTeX Template
%% https://github.com/carlacisternasg/CV-LuaLaTeX

%% LaTeX Project Public License v1.3c
%% https://github.com/bgonzalezbustamante/CV-XeLaTeX/blob/master/LICENSE.md

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cvsection{Relevant Research Projects}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Unpacking the Unpredictable: Using NLP and LLMs to Examine Cabinet Politics and Responses to Stochastic Events in Presidential Democracies}
{Enlace-Inserción Vice-Chancellor Research Office at the Universidad Diego Portales (UDP), {\bfseries PI}, about USD\$4 000. Period: 2025-2026.} 
{UDP}
{2025}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Making Finance Sustainable? A Comparative Perspective on the Politics of Investment}
{NWO VIDI SWG Project VI.Vidi.211.226, {\bfseries PostDoc} (Natascha van der Zwan, PI), about USD\$1 000 000. Period: 2024-2026.} %% {\pounds} 800 000
{Dutch Research Agency}
{2024}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Validación de un estándar de referencia para medir toxicidad e incivilidad política en la esfera digital}
{Internal Project 2024 at the Faculty of Administration and Economics at the Universidad Diego Portales (UDP), {\bfseries PI}, about USD\$2 000. Period: 2024-2025.} %% {\pounds} 1 500
{UDP}
{2024}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Large Language Models (LLMs) to Identify Toxicity in the Digital Sphere during Protest Events in Latin America}
{OpenAI Academic Programme, {\bfseries PI}, about USD\$5 000. Period: 2024.}
{OpenAI}
{2024}
\end{cvhonors}

\vspace{1mm}

\begin{cvhonors}
\cvhonor
{Unravelling the Rule of Law: A Machine Learning Approach}
{NWO XS Project 406.XS.01.013, {\bfseries PostDoc} (Jaroslaw Kantorowicz, PI), about USD\$60 000. Period: 2023-2024.} %% {\pounds} 45 000
{Dutch Research Agency}
{2023}
\end{cvhonors}

%% \vspace{1mm}

%% \begin{cvhonors}
%% \cvhonor
%% {Patrones territoriales de gobierno electrónico y abierto en los municipios chilenos}
%% {Internal Project USA 2156 POC2022\_FAE2 at the Universidad de Santiago de Chile (USACH), {\bfseries PI}, about USD\$2 250. Period: 2022.} %% {\pounds} 1 700
%% {USACH}
%% {2022}
%% \end{cvhonors}

%% \vspace{1mm}

%% \begin{cvhonors}
%% \cvhonor
%% {Sistema político y visibilidad de actores en el debate público articulado por la prensa de élite: 1990-2018}
%% {XVI Internal Competition for the Promotion of Research 2019 at the Universidad Alberto Hurtado (UAH), {\bfseries Research Associate} (Ximena Orchard, PI), about USD\$4 000. Period: 2019.} %% {\pounds} 3 000
{%% UAH}
%% {2019}
%% \end{cvhonors}

\vspace{1mm}
%% \pagebreak